% file : bkfzhead.tex
\markboth{}{}             % empty foot
\vspace*{20pt}
\begin{center} {\Large\bf ZEBRA version 3.66} \end{center}
\vspace*{30pt}
\begin{description}
  \item[Chapter 1] User specifications for the FZ package
  \begin{itemize}
     \item[1.01] Representation of a data-structure
     \item[1.02] Events, Runs, and Files
     \item[1.03] Outline of usage for sequential medium 'tape' or 'disk'
     \item[1.04] FZFILE - initialize a ZEBRA file
     \item[1.05] FZLOGL - change the logging level of a file
     \item[1.06] FZMEMO - connect user memory area for medium 'memory'
     \item[1.07] FZHOOK - connect user routine for medium 'channel'
     \item[1.08] FZLIMI - limit the size of an output file
     \item[1.09] FZODAT - storing and recovering the Direct Access Table
     \item[1.10] FZRUN  - write a RUN record
     \item[1.11] FZOUT  - write one data-structure
     \item[1.12] FZIN   - read one data-structure
     \item[1.13] FZIN   - read one data-structure by segments
     \item[1.14] FZINXT - reset the read point on a direct access file
     \item[1.15] FZCOPY - copy one data-structure from input to output
     \item[1.16] FZENDO - output file termination
     \item[1.17] FZENDI - input file termination
     \item[1.18] Usage for medium 'memory'
     \item[1.19] Usage for medium 'channel'
     \item[1.20] Usage for random access
     \item[1.21] User marking of data-structures for FZOUT
     \item[1.22] Suppress loading of unused parts of FZ
     \item[1.23] FZ installation options
  \end{itemize}
  \item[Chapter 2] Machine dependent usage of FZ files
  \item[Chapter 3] Format specifications
  \begin{itemize}
     \item[3.01] Native mode
     \item[3.02] Exchange mode
     \item[3.03] Exchange Data format
     \item[3.04] Exchange File Format
     \item[3.05] Example for coding dedicated on-line output
  \end{itemize}
  \item[Chapter 4] Technical details
\end{description}
\vspace*{10pt}
\begin{center}
  J.Zoll\\
  December 1990
\end{center}
\cleardoublepage  % continue on next odd page
\begin{center}
  {\large\bf{Principles}}
\end{center}
\par
FZOUT transfers a
\underline{data-structure}
from the dynamic store to a sequential file, i.e. to disk or tape,
or to flat memory, or via a user routine to the outside world.
FZIN does the inverse transfer.
In high-energy physics most FZ files contain data-structures
representing events; but the concepts are completely general.
\par
Two kinds of user information can be associated to a
data-structure on the file.
The 'user header vector' may contain identifying information,
essentially numeric, to allow rapid selection of data-structures.
The 'text vector' may contain zero,
one, or several lines of Hollerith text,
giving context information about the data-structure in one of the
natural languages
(the implementation of text vectors is pending).
\par
On the file, a data-structure is represented by a straight
copy of the relevant sections of the dynamic store,
preceded by its 'pilot information' which describes
the data-structure to follow.
It predicts the memory requirements of the d/s,
and it carries the relocation table necessary to recalculate all
links in the data-structure for the new position in memory.
It further contains the associated user header vector
and the text vector.
Files to be read back on the computer of creation can be written in
'\underline{native mode}'.
In this mode, writing and reading is done with standard FORTRAN
unformatted WRITE and READ statements,
the representation of numbers and Hollerith is defined by the
internal representation of the machine,
blocking of logical records into physical records is defined by the
FORTRAN I/O system of the machine.
The data are written directly from the dynamic store to the file
without going through a Zebra buffer;
a data-structure is hence represented in general by several
logical records.
\par
Files to be transported from one computer to another one of a
different make must be written in
'\underline{exchange mode}'.
Exchange mode implies two separate features :
\begin{itemize}
 \item[--] The Zebra Exchange \underline{File Format}
 provides for reading and writing machine independent files :
 it supports logical records blocked
 onto a string of fixed-length physical records.
 The data flow through a Zebra buffer,
 and each data-structure written by FZOUT is one single logical record.
 For transport of the data via a network which does not allow
 transmission of binary files, the 'ALFA exchange format' is available :
 this maps each physical record onto a string of 80 column card-images
 containing only alpha-numeric characters (and a few others).
 Such a file can be sent as ordinary text over any network;
 it is even possible to include some test events onto the
 PAM file of some program.
 \item[--] The Zebra Exchange \underline{Data Format}
 provides a machine independent interface for the representation
 of the data.
 On each machine Zebra is capable to convert between the
 machine internal and the exchange representation;
 this relies on the I/O characteristic carried by each bank describing
 the nature of its contents (integer, floating, etc.).
 On most modern 32-bit machines the 'native' data format is identical
 to the exchange representation;
 thus on these machines no conversion is needed.
 \item[--] It is possible to combine the exchange file format
 with the native data format;
 this permits to parcel the data into fixed length records,
 without also translating to or from the exchange data
 representation.
\end{itemize}
\par
If the file medium is
\underline{Disk} or \underline{Tape}
the records representing a data-structure are transfered
between the Zebra store and the medium by WRITE or READ statements
(or equivalent).
If the 'file' medium is
\underline{Memory}
the records are tranfered by copying to or from a region
of the user's memory
(possibly involving packing or byte inversion operations).
If the file medium is
\underline{Channel}
the records are handed to, or obtained from,
a user routine one at a time.
This routine is supposed to be an interface to a computer link,
permitting to shuffle the records of the data-structure
from one machine to another one.
Thus one may have FZOUT on one machine feeding FZIN on
the other machine in real time.
The same can be achieved using the medium Memory,
but in this case one needs an area of user memory large enough
to hold the largest d/s.
For the media Memory and Channel FZIN/FZOUT always operate
with the Exchange File Format,
ie. with a string of fixed length records,
but the Data Format may be either 'exchange' or 'native'.
\par
A disk file with Exchange File Format can be read with
\underline{Direct Access}
permitting a random access to the data, since it is a
simple string of fixed-length records.
For each data-structure written (or read) FZOUT or FZIN deliver
its D/A address which the user can compile into a Direct Access Table
together with other relevant information about each d/s.
Handing a D/A address to FZINXT will reset the current read point
of the file, and the next call to FZIN will deliver the wanted d/s.
If the Direct Access table is put into a bank it can be stored
at the end of the file itself by calling FZODAT,
to be recovered by FZIDAT when one comes back to read the file.
\par
The Fortran implementations on some machines running Unix
cannot handle sequential access of fixed-length records,
they insist on having some system control words with each record.
One can get around this by using the Direct Access mechanisme :
Zebra operates sequentially, also for a file which has been
initialized for direct access, until the user tells it otherwise
by calling FZINXT.
However, this is not a satisfactory solution for handling tapes.
Therefore, yet another mode, reading and writing through calls
to the
\underline{C Library}, is programmed into Zebra, both for sequential
and for direct access, but this only on some machines running
the Unix operating system.
\par
It was one of the design aims of the FZ package to provide
a representation of the data on external media in exchange mode
which fits both off-line data-processing and on-line data-taking
requirements.
Speed is important in both cases, but it is more critical in the
on-line context.
For this the ZEBRA Exchange File Format has been defined to allow
dumping large areas of memory to a file,
without the need  to insert control information  for physical
records during the dumping process;
all control information needed is grouped at the very beginning
of the data.
\par
The program running on a data-taking on-line computer is
likely not to be a ZEBRA program,
in which case one will have to write an ad hoc output routine to
produce a file readable by FZIN in exchange mode.
This point has been kept in mind when designing the exchange format,
to make sure that it is simple enough
to produce files in this format.
\par
Although the unit of information for ZEBRA is a data-structure
in full generality,
in the on-line application the
'data-structure' will most likely consist of just one maybe
very large bank,
or at most of a few banks.
For such simple data-structures the interconnection
by links does not need to exist,
and the on-line program can disregard this aspect of
data-structures.
When a 'link-less' data-structure of several banks is read
all its banks will be connected by FZIN into
a linear structure, to permit the reading program to access
the banks.
