\Version{ERF}         \Routid{C300}
\Keywords{ERROR COMPLEMENTARY FUNCTION}
\Author{G.A. Erskine} \Library{MATHLIB or Fortran Compiler Library}
\Submitter{K.S. K\"olbig}         \Submitted{20.04.1970}
\Language{Fortran}    \Revised{07.06.1992}
\Cernhead{Error Function and Complementary Error Function}
Function subprograms {\tt ERF}, {\tt ERFC} and {\tt DERF}, {\tt DERFC}
compute the error and complementary error functions
$$ \mathrm{erf}(x)  \ = \
\displaystyle \frac{2}{\sqrt\pi}\int_0^x  e^{-t^2}\, dt, \qquad
\mathrm{erfc}(x) \ = \
\displaystyle \frac{2}{\sqrt \pi}\int_x^\infty e^{-t^2}\, dt, $$
defined for all values of the real argument $x$.
\par
On CDC and Cray computers, the double-precision versions
{\tt DERF} and {\tt DERFC} are not available.
\Structure
{\tt FUNCTION} subprograms\\
User Entry Names: \Rdef{ERF}, \Rdef{ERFC}, \Rdef{DERF}, \Rdef{DERFC}
\Usage
In any arithmetic expression,
\begin{center}
\parbox{.6\textwidth}{
{\tt ERF(X) } \quad or \quad {\tt DERF(X) }
\quad has the value \quad erf({\tt X}),\\
{\tt ERFC(X)} \quad or \quad {\tt DERFC(X)}
\quad has the value \quad
erfc({\tt X}),}
\end{center}
where {\tt ERF}, {\tt ERFC}, are of type {\tt REAL},
{\tt DERF}, {\tt DERFC}, are of type
{\tt DOUBLE PRECISION}, and {\tt X} has the same type as the function
name.
\Method
Computation by rational Chebyshev approximation.
\Accuracy
The system-supplied versions (see {\bf Notes}) have full machine
accuracy. The CERN-supplied versions of {\tt ERF} and {\tt ERFC} have
full single-precision accuracy (slightly less on CDC and Cray
computers). The CERN-supplied versions of {\tt DERF} and {\tt DERFC}
have an accuracy  of 15 significant digits.
\Notes
On some computers,  one or both of these functions
is  available in the  system-supplied Fortran
mathematical  library. In this case the system-supplied
version will be loaded  instead of the CERN version.
\Refer
\begin{enumerate}
\item W.J. Cody, Rational Chebyshev approximations for the error
function, Math. Comp. {\bf 22} (1969) 631--637.
\end{enumerate}
$\bullet$
