%  12 jan 95  ksk
\Version {ALGAMA}                            \Routid{C304}
\Keywords{LOGARITHM GAMMA FUNCTION}
\Author{K.S. K\"olbig} \Library{MATHLIB or Fortran Compiler Library}
\Submitter{}                             \Submitted{07.06.1992}
\Language{Fortran}                       \Revised{15.03.1993}
\Cernhead {Logarithm of the Gamma Function}
Function subprograms {\tt ALGAMA}, {\tt DLGAMA} and {\tt QLGAMA}
compute the logarithm of the gamma function
$$\displaystyle \ln \Gamma(x) \ = \
\ln \int_0^\infty e^{-t} t^{x-1}\, dt \qquad (x>0) $$
for real argument $x>0$.
\par
The quadruple-precision version {\tt QLGAMA} is available only on
computers which support a {\tt REAL*16} Fortran data type.
\Structure
{\tt FUNCTION} subprograms \\
User Entry Names: \Rdef{ALGAMA}, \Rdef{DLGAMA}, \Rdef{QLGAMA}  \\
Obsolete User Entry Names: \Rdef{ALOGAM} $\equiv$ {\tt ALGAMA},
\Rdef{DLOGAM} $\equiv$ {\tt DLGAMA} \\
Files Referenced: {\tt Unit 6} \\
External References: \Rind {MTLMTR}{N002}, \Rind{ABEND}{Z035}
\Usage
In any arithmetic expression,
\begin{center}
{\tt ALGAMA(X)}, \quad {\tt DLGAMA} \quad or \quad {\tt QLGAMA(X)}
\quad has the value \quad $\ln \Gamma(\mathtt{X})$,
\end{center}
where {\tt ALGAMA} is of type {\tt REAL}, {\tt DLGAMA} is of type
{\tt DOUBLE PRECISION}, {\tt QLGAMA} is of type {\tt REAL*16},
and {\tt X} has the same type as the function name.
\Method
Rational approximations.
\Accuracy
The system-supplied version (see {\bf Notes}) has full machine
accuracy.
The CERN-supplied version of {\tt ALGAMA} (except on CDC and Cray
computers) has full single-precision accuracy.
For most values of the argument {\tt X},
the CERN-supplied versions of {\tt DLGAMA}, {\tt QLGAMA}
(and of {\tt ALGAMA}, {\tt DLGAMA}
on CDC and Cray computers) have an accuracy of approximately one
significant digit less than the machine precision.
\Errorh
Error {\tt C304.1:}  $\mathtt{X \le 0}$.
The function value is set equal to zero, and a message is written on
on {\tt Unit 6}, unless subroutine {\tt MTLSET} (N002) has been called.
\Notes
If the function {\tt ALGAMA} or {\tt DLGAMA} is available in the
system-supplied Fortran mathematical library, the system-supplied
function will be loaded instead of the CERN version.
\Refer
\begin{enumerate}
\item  W.J. Cody and K.E. Hillstrom, Chebyshev approximations
for the natural logarithm of the gamma function,
Math. Comp. {\bf 21} (1967) 198--203.
\item  J.F. Hart et al., Computer approximations (John Wiley
 Sons, New York 1968) 287.
\end{enumerate}
$\bullet$
