\Version{CGPLG}                  \Routid{C321}
\Keywords{GENERAL NIELSEN POLYLOGARITHM}
\Author{K.S. K\"olbig}            \Library{MATHLIB}
\Submitter{}                       \Submitted{12.09.1985}
\Language{Fortran}               \Revised{15.03.1993}
\Cernhead{Nielsen's Generalized Polylogarithm}
Function subprograms {\tt CGPLG} and {\tt WGPLG}
calculate the complex-valued generalized polylogarithm function
$$ S_{n,m}(x) \ = \ \frac{(-1)^{n+m-1}}{(n-1)!\,m!}\,\int^1_0
t^{-1}\ln^{n-1} t \,\ln^m (1-xt)\,dt \qquad (*) $$
for real arguments $x$ and integer $n$ and $m$ satisfying
$1\leq n\leq 4,\  1\leq m\leq 4,\  n+m\leq 5$;
i.e., one of the functions
$S_{1,1}$, $ S_{1,2}$, $ S_{2,1}$, $ S_{1,3}$,
$S_{2,2}$, $ S_{3,1}$, $ S_{1,4}$, $ S_{2,3}$,
$S_{3,2}$, $ S_{4,1}$. If $x\leq1$, $S_{n,m}(x)$ is real,
and the imaginary part is set equal to zero.
\par
The double-precision version {\tt WGPLG} is available only on computers
which support a {\tt COMPLEX*16} Fortran data type.
\Structure
{\tt FUNCTION} subprograms\\
User Entry Names: \Rdef{CGPLG}, \Rdef{WGPLG}\\
Files Referenced: {\tt Unit 6} \\
External References: \Rind{MTLMTR}{N002}, \Rind{ABEND}{Z035}
\Usage
In any arithmetic expression,
\begin{center}
{\tt CGPLG(N,M,X)} \quad or \quad {\tt WGPLG(N,M,X)} \quad has the value
\quad $S_{\mathtt{N,M}}(\mathtt{X})$,
\end{center}
where {\tt CGPLG} is of type {\tt COMPLEX}, {\tt WGPLG} is of type
{\tt COMPLEX*16}, {\tt X} is of type {\tt REAL}
for {\tt CGPLG} and of type {\tt DOUBLE PRECISION} for {\tt WGPLG},
and where {\tt N} and {\tt M} are of type {\tt INTEGER}.
\Method
The method is described in Ref. 1.
Note that the imaginary part of the function defined as $S_{n,m}(x)$
in Ref. 1 has the opposite sign to the
imaginary part of the function defined by (*). See Ref. 2.
\Accuracy
{\tt CGPLG} (except on CDC and Cray computers)
has full single-precision accuracy.
For most values of the argument {\tt X}, {\tt WGPLG}
(and {\tt CGPLG} on CDC and Cray computers) has an accuracy of
approximately two significant digits less than the machine precision.
The loss of accuracy is greater when {\tt X} is very close to {\tt 1}.
\Errorh
Error {\tt C321.1}: $\mathtt{N,M < 1}$ or $\mathtt{N,M > 4}$ or
$\mathtt{N+M > 5}$.
The function value is set equal to zero, and a message is written on
{\tt Unit 6}, unless subroutine {\tt MTLSET} (N002) has been called.
\Refer
\begin{enumerate}
\item K.S. K\"olbig, J.A. Mignaco and
E. Remiddi, On Nielsen's generalized polylogarithms and their
numerical calculation, BIT {\bf 10} (1970) 38--71.
\item K.S. K\"olbig, Nielsen's generalized polylogarithms,
SIAM J. Math. Anal. {\bf 17} (1986) 1232--1258.
\end{enumerate}
$\bullet$
