% 31 oct 94  ksk
\Version{RFERDR}                       \Routid{C323}
\Keywords{FERMI DIRAC FUNCTION}
\Author{K.S. K\"olbig}                \Library{MATHLIB}
\Submitter{}                          \Submitted{15.05.1987}
\Language{Fortran}                    \Revised{01.12.1994}
\Cernhead{Fermi-Dirac Function}
Function subprograms {\tt RFERDR} and {\tt DFERDR} calculate
the Fermi-Dirac function
$$ F_k(x) \ = \ \int^\infty_0\frac{t^{k/2}}{1+e^{t-x}}\,dt $$
for real argument $x$, and $k=-1,1,3.$
\par
On CDC and Cray computers, the double-precision version
{\tt DFERDR} is not available.
\Structure
{\tt FUNCTION} subprograms\\
User Entry Names: \Rdef{RFERDR}, \Rdef{DFERDR} \\
Obsolete User Entry Names: \Rdef{FERDR} $\equiv$ \Rdef{RFERDR} \\
External References: \Rind{MTLMTR}{N002}, \Rind{ABEND}{Z035}
\Usage
In any arithmetic expression,
\begin{center}
{\tt RFERDR(X,K)} \quad or \quad {\tt DFERDR(X,K)} \quad has the value
\quad $F_{\mathtt{K}}(\mathtt{X})$,
\end{center}
where {\tt RFERDR} is of type  {\tt REAL,} {\tt DFERDR} is of type
{\tt DOUBLE PRECISION}, and {\tt X} has the same type as the
function name. {\tt K (INTEGER)} = {\tt -1}, or {\tt 1} or {\tt 3}.
\Method
Rational approximation.
\Accuracy
{\tt RFERDR} (except on CDC and Cray computers)
has full single-precision accuracy.
For most values of the argument {\tt X}, {\tt DFERDR}
(and {\tt RFERDR} on CDC and Cray computers) has, for $\mathtt{X>0}$,
an accuracy of 7-10 digits and for $\mathtt{X<0}$, an accuracy of
10 to 14 digits.
\Errorh
Error {\tt C323.1}: $\mathtt{K \ne -1,1,3}.$
The function value is set equal to zero, and a message is written on
{\tt Unit 6}, unless subroutine {\tt MTLSET} (N002) has been called.
\Refer
\begin{enumerate}
\item W.J. Cody and H.C. Thacher,Jr., Rational approximations for
Fermi-Dirac integrals of order $-1/2$, $1/2$ and $3/2$,
Math. Comp. {\bf 21} (1967) 30--40.
\end{enumerate}
$\bullet$
