%  09 may 96  ksk
\Version{RCLAUS}                         \Routid{C326}
\Keywords{CLAUSEN FUNCTION}
\Author{K.S. K\"olbig}                   \Library{MATHLIB}
\Submitter{}                             \Submitted{01.12.1994}
\Language{Fortran}                   %    \Revised{}
\Cernhead{Clausen Function}
Function subprograms {\tt RCLAUS} and {\tt DCLAUS} calculate the
Clausen function
$$ \mathrm{Cl}_2(x) \ = \ \displaystyle
- \int_0^x \ln \left|2 \sin \frac{t}{2}\right|\,dt \ = \
\sum_{k=1}^\infty \frac{\sin kx}{k^2} $$
for real arguments $x$.
\par
On CDC and Cray computers, the double-precision version
{\tt DCLAUS} is not available.
\Structure
{\tt FUNCTION} subprograms\\
User Entry Names: \Rdef{RCLAUS}, \Rdef{DCLAUS}
\Usage
In any arithmetic expression,
\begin{center}
{\tt RCLAUS(X)} \quad or \quad {\tt DCLAUS(X)} \quad has the value \quad
$\mathrm{Cl}_2(\mathtt{X})$,
\end{center}
where {\tt RCLAUS} is of type {\tt REAL}, {\tt DCLAUS} is of type
{\tt DOUBLE PRECISION}, and {\tt X} has the same type as the function
name.
\Method
For $0 \le x \le \pi$, the function is approximated by truncated
Chebyshev series. For $x$ outside this range, the relations
$\mathrm{Cl}_2(\pi+x)=-\mathrm{Cl}_2(\pi-x)$ and
$\mathrm{Cl}_2(2n\pi \pm x)=\pm \mathrm{Cl}_2(x)$ are used.
\Accuracy
{\tt RCLAUS} (except on CDC and Cray computers)
has full single-precision accuracy in the interval $0 \le x \le 2\pi$.
For most values of the argument $\mathtt{X} \in [0,2\pi]$, {\tt DCLAUS}
(and {\tt RCLAUS} on CDC and Cray computers) has an accuracy of
approximately one significant digit less than the machine precision.
Accuracy is lost near the zero of $\mathrm{Cl}_2(x)$ at $x=\pi$
and for large values of $|x|$.
\Refer
\begin{enumerate}
\item K.S. K\"olbig, Chebyshev coefficients for the Clausen function
$\mathrm{Cl}_2(x)$, J. Comput. Appl. Math. {\bf 64} (1995) 295--297.
\end{enumerate}
$\bullet$
