% 08 jun 1995 ksk
\Version{RELI1}                       \Routid{C346}
\Keywords{ELLIPTIC INTEGRAL FIRST SECOND THIRD KIND INCOMPLETE}
\Author{K.S. K\"olbig}                  \Library{MATHLIB}
\Submitter{ }                           \Submitted{07.06.1992}
\Language{Fortran}                      \Revised{}
\Cernhead{Elliptic Integrals of First, Second, and Third Kind}
Function subprograms {\tt RELI1}, {\tt RELI2}, {\tt RELI3}
and {\tt DELI1}, {\tt DELI2}, {\tt DELI3} calculate, for real
argument $x$, the elliptic integrals of the first, second and
third kind, respectively.
\par
On CDC and Cray computers, the double-precision versions {\tt DELI1},
{\tt DELI2} and {\tt DELI3} are not available.
\par
Mainly for reasons of numerical stability, the algorithms are based on
the following definitions:
\par
{\bf First kind:}
\begin{eqnarray*}
\mathbf{F}_1(x,k') \ = \  \displaystyle
\int_0^x \frac{d\xi}{\sqrt{(1+\xi^2)(1+{k'}^2\xi^2)}}
\qquad ({k'}^2 \ge 0).
\end{eqnarray*}
{\bf Second kind:}
\begin{eqnarray*}
\mathbf{F}_2(x,k',a,b) \ = \ \displaystyle
\int_0^x \frac{a+b\xi^2}{(1+\xi^2)\sqrt{(1+\xi^2)(1+{k'}^2\xi^2)}}\,d\xi
\qquad ({k'}^2 \ge 0).
\end{eqnarray*}
{\bf Third kind:}
\begin{eqnarray*}
\mathbf{F}_3(x,k',p) \ = \ \displaystyle
\int_0^x \frac{1+\xi^2}{(1+p\xi^2)\sqrt{(1+\xi^2)(1+{k'}^2\xi^2)}}\,d\xi
\qquad ({k'}^2 \ge 0, \ px^2 \ne -1).
\end{eqnarray*}
\par
Note that $\mathbf{F}_1(x,k') = \mathbf{F}_2(x,k',1,1) =
\mathbf{F}_3(x,k',1)$.
For $p < 0$, the integral $\mathbf{F}_3$ is defined by its principal
value.
\par
For the integral of the second kind, a special entry-mode argument
is provided which allows $\mathbf{F}_2(x,k',a,b)$ to be calculated when
${k'}^2 < 0$, i.e. when $k'$ is imaginary.
\par
Other common definitions of the elliptic integrals and their
relations to $\mathbf{F}_1$, $\mathbf{F}_2$, $\mathbf{F}_3$ are
listed here for convenience ($k^2+{k'}^2=1$):
\par
{\bf First kind:}
\begin{eqnarray*}
F(k,\phi) \ = \ \displaystyle \int_0^\phi
\frac{d\psi}{\sqrt{1-k^2\sin^2 \psi}} \ = \ \mathbf{F}_1(\tan \phi,k')
\qquad (|\phi| \le \pi/2, \ |k| < 1),
\end{eqnarray*}
\begin{eqnarray*}
\widehat{F}(y,k) \ = \ \displaystyle \int_0^y
\frac{d\eta}{\sqrt{(1-\eta^2)(1-k^2\eta^2)}} \ = \
\mathbf{F}_1 \left( y/\sqrt{1-y^2},k' \right)
\qquad (|y| < 1, \ |k| < 1).
\end{eqnarray*}
{\bf Second kind:}
\begin{eqnarray*}
E(k,\phi) \ = \ \displaystyle \int_0^\phi
\sqrt{1-k^2\sin^2 \psi} \, d\psi \ = \
\mathbf{F}_2(\tan \phi,k',1,{k'}^2)
\qquad (|\phi| \le \pi/2, \ |k| \le 1),
\end{eqnarray*}
\begin{eqnarray*}
\widehat{E}(y,k) \ = \ \displaystyle \int_0^y
\sqrt{\frac{1-k^2 \eta^2}{1-\eta^2}} \, d\eta \ = \
\mathbf{F}_2 \left( y/\sqrt{1-y^2},k',1,{k'}^2 \right)
\qquad (|y| < 1, \ |k| \le 1).
\end{eqnarray*}
\newpage
{\bf Third kind:}
$$ \begin{array}{rclcl}
\Pi(\phi,h,k) & = & \displaystyle \int_0^\phi
\frac{d\psi}{(1+h\sin^2 \psi)\sqrt{1-k^2\sin^2 \psi}} & = &
\mathbf{F}_3(\tan \phi,k',h+1) \\
& & & & (|\phi| \le \pi/2, \ |k| < 1), \\[3mm]
\widehat{\Pi}(y,h,k) & = & \displaystyle \int_0^y
\frac{d\eta}{(1+h\eta^2)\sqrt{(1-\eta^2)(1-k^2\eta^2)}} & = &
\mathbf{F}_3 \left( y/\sqrt{1-y^2},k',h+1 \right) \\
& & & & (|y| < 1, \ |k| < 1).
\end{array} $$
\Structure
{\tt FUNCTION} subprograms \\
User Entry Names:
\Rdef{RELI1}, \Rdef{RELI2}, \Rdef{RELI3}, \Rdef{DELI1},
\Rdef{DELI2}, \Rdef{DELI3} \\
Files Referenced: {\tt Unit 6} \\
External References: \Rind{ASINH}{B102}, \Rind{DASINH}{B102},
\Rind{MTLMTR}{N002}, \Rind{ABEND}{Z035}
\Usage
In any arithmetic expression, with $\mathtt{AKP}=k'$,
\begin{center}
\begin{tabular}{l@{\quad or \quad}l@{\quad has the value \quad}l}
{\tt RELI1(X,AKP)} & {\tt DELI1(X,AKP)} &
$\mathbf{F}_1(\mathtt{X},k')$,\\
{\tt RELI2(X,AKP,A,B,MODE)} & {\tt DELI2(X,AKP,A,B,MODE)} &
$\mathbf{F}_2(\mathtt{X},k'',\mathtt{A,B})$,\\
{\tt RELI3(X,AKP,P)} & {\tt DELI3(X,AKP,P)} &
$\mathbf{F}_3(\mathtt{X},k',\mathtt{P})$,\\
\end {tabular}
\end{center}
where {\tt RELI1}, {\tt RELI2}, {\tt RELI3} are of type {\tt REAL},
where {\tt DELI1}, {\tt DELI2}, {\tt DELI3} are of type
{\tt DOUBLE PRECISION}, and {\tt X}, {\tt AKP}, {\tt A}, {\tt B} and
{\tt P} have the same type as the function name. {\tt MODE} is of type
{\tt INTEGER}.
\par
The notation $k''$ indicates that, when calling {\tt RELI2} or
{\tt DELI2}, the parameters {\tt AKP} and {\tt MODE} must be set
as follows: \\
If ${k'}^2 > 0$: \ $\mathtt{MODE = +1}$ and $\mathtt{AKP} = k'$, \\
if ${k'}^2 < 0$: \ $\mathtt{MODE = -1}$ and
$\mathtt{AKP} =$ Im $k' = -ik'$ (real).
\Method
The evaluation of $\mathbf{F}_1$ and $\mathbf{F}_2$ is based
on the Landen
transformation, that of $\mathbf{F}_3$ on the Bartky transformation.
$\mathbf{F}_2$ for ${k'}^2 < 0$ is calculated by using a transformation of
the arguments. See Ref. 1 and 2 for details.
\Accuracy
The {\tt REAL} functions (except on CDC and Cray computers) have full
single-precision accuracy. The {\tt REAL} functions on CDC and Cray, and
the {\tt DOUBLE PRECISION} functions on all computers have an accuracy
approximately two significant digits less than the machine precision.
\Restrict
1. \quad {\tt RELI2} and {\tt DELI2:} \
{\tt AKP*X**2 $<$ 1} if {\tt MODE $=-$1}.  \\
2. \quad {\tt RELI2} and {\tt DELI2:} \
$\mathtt{MODE = +1}$ or $\mathtt{-1}$. \\
3. \quad {\tt RELI3} and {\tt DELI3:} \ {\tt P*X**2 $\ne -$1}.
\Errorh
Error {\tt C346.1:} Restriction 1 is not satisfied. \\
Error {\tt C346.2:} Restriction 2 is not satisfied. \\
Error {\tt C346.3:} Restriction 3 is not satisfied. \\
In all cases, the function value
is set equal to zero, and a message is written on {\tt Unit 6},
unless subroutine {\tt MTLSET} (N002) has been called.
\newpage
\Source
The subprograms are based on the Algol60 procedures {\it el1, el2}
in Ref. 1 and {\it el3} in Ref. 2.
\Refer
\begin{enumerate}
\item R. Bulirsch, Numerical calculation of elliptic integrals and
elliptic functions, Numer. Math. {\bf 7} (1965) 78--90.
\item R. Bulirsch, Numerical calculation of elliptic integrals and
elliptic functions III, Numer. Math. {\bf 13} (1969) 305--315.
\end{enumerate}
$\bullet$
