% 5 june 1992 mg
\Version{CELINT}                       \Routid{C348}
\Keywords{ELLIPTIC INTEGRAL COMPLEX GENERAL FIRST SECOND KIND}
\Keywords{INCOMPLETE}
\Author{K.S. K\"olbig}                  \Library{MATHLIB}
\Submitter{ }                           \Submitted{07.06.1992}
\Language{Fortran}                      \Revised{}
\Cernhead{Elliptic Integral for Complex Argument}
Function subprograms {\tt CELINT} and {\tt WELINT} calculate, for complex
argument $z=x+iy$ and real complementary modulus $k'$ a general
elliptic integral of the second kind:
$$ \mathbf{F}(z,k',a,b) \ = \ \displaystyle
\int_0^z \frac{a+b\zeta^2}{(1+\zeta^2)
\sqrt{(1+\zeta^2)(1+{k'}^2\zeta^2)}} \, d\zeta
\qquad ({k'}^2 \geq 0, \ \mathrm{Re}(z) \geq 0), $$
which contains the elliptic integrals of the first and second kind as
special cases:
$$ \begin{array}{rclcl}
\mathbf{F}_1(z,k') & = & \displaystyle \int_0^z \frac{d\zeta}
{\sqrt{(1+\zeta^2)(1+{k'}^2 \zeta^2)}} & = & \mathbf{F}(z,k',1,1),\\[6mm]
\mathbf{F}_2(z,k') & = & \displaystyle \int_0^z \frac{d\zeta}
{(1+\zeta)^2} \sqrt{\frac{1+{k'}^2 \zeta^2}{1+\zeta^2}} & = &
\mathbf{F}(z,k',1,{k'}^2).
\end{array} $$
\par
The double-precision version {\tt WELINT} is available only on computers
which support a {\tt COMPLEX*16} Fortran data type.
\Structure
{\tt FUNCTION} subprograms \\
User Entry Names: \Rdef{CELINT}, \Rdef{WELINT} \\
Files Referenced: {\tt Unit 6} \\
External References: \Rind{MTLMTR}{N002}, \Rind{ABEND}{Z035}
\Usage
In any arithmetic expression, with $\mathtt{AKP}=k'$,
\begin{center}
{\tt CELINT(Z,AKP,A,B)} \quad or \quad {\tt WELINT(Z,AKP,A,B)} \quad
has the value \quad $\mathbf{F}(\mathtt{Z},k',\mathtt{A,B})$,
\end{center}
where {\tt CELINT} is of type {\tt COMPLEX}, {\tt WELINT} is of type
{\tt COMPLEX*16}, {\tt Z} is of the same type as the function name,
and {\tt AKP}, {\tt A}, {\tt B} are of type {\tt REAL} for {\tt CELINT}
and of type {\tt DOUBLE PRECISION} for {\tt WELINT}.
\Method
The evaluation of $\mathbf{F}$ is based on the Gauss transformation.
For details, in particular for the conformal mapping provided by
$\mathbf{F}$, see Ref. 1.
\Accuracy
{\tt CELINT} (except on CDC and Cray computers)
has full single-precision accuracy.
For most values of the arguments, {\tt WELINT}
(and {\tt CELINT} on CDC and Cray computers) has an accuracy of
approximately one significant digit less than the machine precision.
\Errorh
Error {\tt C348.1:} Re $\mathtt{Z < 0}$.
The function value is set equal to zero, and a message is written on
{\tt Unit 6}, unless subroutine {\tt MTLSET} (N002) has been called.
\newpage
\Notes
For other forms of the elliptic integrals see the write-up
for {\tt RELI1} (C346).
\Source
The subprogram is based on the Algol60 procedure {\it elco2}
given in Ref. 1.
\Refer
\begin{enumerate}
\item R. Bulirsch, Numerical calculation of elliptic integrals and
elliptic functions, Numer. Math. {\bf 7} (1965) 78--90.
\end{enumerate}
$\bullet$
