% 20 oct 94  ksk
\Version{RDEQBS}                          \Routid{D201}
\Keywords{FIRST-ORDER DIFFERENTIAL EQUATION GRAGG BULIRSCH STOER}
\Authors {K.S. K\"olbig}                 \Library{MATHLIB}
\Submitter{}                             \Submitted{15.02.1989}
\Language{Fortran}                   \Revised{01.12.1994}
\Cernhead{First-order Differential Equations (Gragg--Bulirsch--Stoer)}
Subroutine subprograms {\tt RDEQBS}  and {\tt DDEQBS} advance the
solution of the system of $n \ge 1$ simultaneous first-order
differential equations
$$ \frac{dy_i}{dx} \ = \
f_i(x,y_1,\ldots,y_n), \qquad (i = 1,2,\ldots,n),$$
from a specified value $x_1$ to a specified value $x_2$ of the
independent variable $x$.
\par
On computers other than CDC and Cray, only the double-precision version
{\tt DDEQBS} is available. On CDC and Cray computers,
only the single-precision version {\tt RDEQBS} is available.
\Structure
{\tt SUBROUTINE} subprograms \\
User Entry Names : \Rdef{RDEQBS}, \Rdef{DDEQBS}\\
Obsolete User Entry Names: \Rdef{DEQBS} $\equiv$ {\tt RDEQBS} \\
Files  Referenced : {\tt Unit 6} \\
External References: \Rind{MTLMTR}{N002}, \Rind{ABEND}{Z035},
user-supplied {\tt SUBROUTINE} subprogram
\Usage
For $\mathtt{t=R}$ (type {\tt REAL}), $\mathtt{t=D}$ (type
{\tt DOUBLE PRECISION}),
\begin{verbatim}
    CALL tDEQBS(N,X1,X2,Y,H0,EPS,SUB,W)
\end{verbatim}
\begin{DLtt}{12345}
\item[N] ({\tt INTEGER}) Number $n$ of equations.
\item[X1] (type according to {\tt t}) Initial value $x_1$ of the
independent variable $x$.
\item[X2] (type according to {\tt t})
Final value $x_2$ of the independent variable $x$.
\item[Y] (type according to {\tt t})
One-dimensional array of length $\ge $ {\tt N}.
On entry, $\mathtt{Y(i),(i=1,\ldots,N)}$, must contain $y_i(x_1)$.
On exit, $\mathtt{Y(i),(i=1,\ldots,N)}$, contains approximate values
 $y_i(x_2)$.
\item[H0] (type according to {\tt t}) On entry, {\tt H0} must contain
the proposed initial step-length $h_0$. On exit, {\tt H0} contains
the last computed step-length (See also {\bf Method} and {\bf Notes}).
\item[EPS] (type according to {\tt t}) The requested absolute
accuracy $\varepsilon$. ({\tt EPS} should not be smaller
than approximately $10^3$ times the machine precision).
\item[SUB] Name of a user-supplied {\tt SUBROUTINE} subprogram,
declared {\tt EXTERNAL} in the calling program.
\item[W] (type according to {\tt t}) Array containing at least
{\tt 36*N} elements required as working-space.
\end{DLtt}
The user-supplied subroutine {\tt SUB} should be of the form
\begin{verbatim}
    SUBROUTINE SUB(X,Y,F)
\end{verbatim}
where the variable {\tt X} and the one-dimensional arrays {\tt Y(*)} and
{\tt F(*)} are of type {\tt t}. This subroutine must set
$$\mathtt{F(I)} = f_{\mathtt{I}}(\mathtt{X,Y(1),\ldots,Y(N))} \qquad
(\mathtt{I = 1,2,\ldots,N}).$$
\newpage
\Method
For the first integration step, starting at $x=x_1$, the step-length
$h$ is chosen to be the smallest of the numbers $h_0,\,h_0/2,\,
h_0/4,\ldots$ for which not more than 9 stages of internal
extrapolation yield an estimated error less than $\varepsilon$. This
procedure is repeated until $x=x_2$ is reached.
(For details, see Ref. 1).
\Errorh
Error {\tt D201.1}: If the requestec accuracy cannot obtained,
a message is written on
{\tt Unit 6}, unless subroutine {\tt MTLSET} (N002) has been called. \\
For $\mathtt{N<1}$, or $\mathtt{X1=X2}$ or $\mathtt{H0=0}$, control
is returned to the calling program without any change in {\tt Y}.
\Notes
For well-conditioned systems of equations any reasonable value of the
initial step length $h_0$ may be chosen. For ill-conditioned systems,
the initial value of $h_0$ may be important, and tests with different
values are advised. An inappropriate choice may lead to wrong results in
such cases.
\Source
This subroutines is based on an Algol60 procedure given in Ref. 1.
The adaption for integration over a given interval (not only over one
step) is due to G. Janin.
\Refer
\begin{enumerate}
\item  R. Bulirsch and J. Stoer, Numerical treatment of ordinary
differential equations by extrapolation methods,
Numer. Math. {\bf 8} (1966) 1--13.
\end{enumerate}
$\bullet$
