\Version {LAPACK}                         \Routid{F001}
\Keywords{linear algebra lapack eispack linpack}
\Author{see below}                       \Library{MATHLIB}
\Submitter{B. Damgaard}                  \Submitted{07.06.1992}
\Language{Fortran}                       \Revised{}
\Cernhead {Linear Algebra Package}
{\bf Authors:} E. Anderson, Z. Bai, C. Bischof, J. Demmel,
 J. Dongarra, J. Du Croz, A. Greenbaum, S. Hammarling, A. McKenney,
  S. Ostrouchov, and D. Sorensen. \\[3mm]
{\tt LAPACK} is
a package of subroutines written in Fortran for solving
the most common problems in numerical linear algebra:
systems of linear equations, linear least squares problems,
eigenvalue problems, and singular value problems.
{\tt LAPACK} is intended to supersede
{\tt LINPACK} and {\tt EISPACK}.
It extends the functionality of these packages by including
equilibration, iterative refinement, error bounds, and driver routines
for linear systems, routines for computing and re-ordering the Schur
factorization, and condition estimation routines for eigenvalue
problems. {\tt LAPACK} improves on the accuracy of the standard
algorithms in {\tt EISPACK} by including high accuracy algorithms for
finding singular values and eigenvalues of bidiagonal and tridiagonal
matrices respectively that arise in SVD and symmetric eigenvalue
problems.
The algorithms and software are structured to achieve high
efficiency on vector processors, high-performance ``superscalar"
workstations, and shared-memory multi-processors.
\Structure
{\tt SUBROUTINE} subprograms
\Usage
It is highly recommended to obtain a copy of the
{\tt LAPACK} Users' Guide published by SIAM.
This Users' Guide gives a detailed description of the
philosophy behind
{\tt LAPACK} as well as an explanation of its usage.
European users must order from the distributors
of SIAM books in Europe:
\begin{quote}
  STM Distribution Ltd.                   \\
  Sunbury International Business Centre   \\
  Middlesex TW16 7DX, England             \\
  Tel. +44 932 765119, \quad FAX +44 932 765429     \\
\end{quote}
or from booksellers.
Other users should contact SIAM directly in order to find out
the address of the local retailer:
\begin{quote}
  SIAM                                    \\
  3600 University City Science Center     \\
  Philadelphia, PA 19104-2688             \\
  Tel. +1 215 382 9800, \quad FAX +1 215 386 7999 \ .         \\
\end{quote}
{\bf   Availability}
 
CERN is
distributing the package only in compiled form,
suited for the CERN-supported platforms. Source code is directly
available via {\tt netlib} (use {\tt find netlib} for details).
Alternatively, NAG offers the distribution via magnetic tapes
for a nominal handling charge. NAG can be contacted at
\begin{quote}
    NAG Response Centre                 \\
    Tel. +44 865 311744, \quad FAX +44 865 311755
\end{quote}
   $\bullet$
