\Version {TR}                          \Routid{F112}
\Keywords{OPERATION SYMMETRIC TRIANGULAR MATRIX}
\Author{W. Hart}                        \Library{KERNLIB}
\Submitter{}                            \Submitted{01.01.1975}
\Language{Fortran}                      \Revised{12.12.1986}
\Cernhead {Manipulation of Triangular and Symmetric Matrices }
At CERN, matrices are often stored row-wise (TC-convention); furthermore,
symmetric matrices are stored packed as the lower left
triangular part only, i.e., the $I$th diagonal element is found in
position $I(I+1)/2$. The {\tt TR}-package performs many of the frequently
required operations associated with such matrices without resorting to
expanding into the unpacked square form. In all the following
routines an $M \times M$  {\it symmetric} matrix is taken to be stored in
the packed form with $M(M+1)/2$ elements.
\par
Some of these operations produce and require the manipulation of
{\it lower triangular} matrices which have all elements zero
above the leading diagonal.
These are also stored in the packed form with all the zeros dropped;
therefore, care has to be taken in the interpretation of a packed matrix
as to whether it represents a symmetric or lower triangular array.
To facilitate this distinction in the Write-up,
the following nomenclature has been adopted:
\begin{DLtt}{12345678}
\item[A,B,C] unpacked rectangular matrices (row-wise storage)
\item[Q,R,S,T] packed symmetric matrices
\item[V,W] packed lower triangular matrices
\end{DLtt}
On 32-bit machines the calculations are performed internally in
double-precision mode.
\Structure
{\tt SUBROUTINE} subprograms  \\
User Entry Names:
\begin{htmlonly}
\begin{tabular}{llllllll}
\end{htmlonly}
\begin{latexonly}
\begin{tabular}[t]{l*{7}{@{\hspace{4pt}}l}}
\end{latexonly}
\Rdef{TRCHUL}, & \Rdef{TRCHLU}, & \Rdef{TRSMUL}, & \Rdef{TRSMLU}, &
\Rdef{TRINV},  & \Rdef{TRSINV}, & \Rdef{TRLA},   & \Rdef{TRLTA},  \\
\Rdef{TRAL},   & \Rdef{TRALT},  & \Rdef{TRSA},   & \Rdef{TRAS},   &
\Rdef{TRSAT},  & \Rdef{TRATS},  & \Rdef{TRAAT},  & \Rdef{TRATA},  \\
\Rdef{TRASAT}, & \Rdef{TRATSA}, & \Rdef{TRQSQ},  & \Rdef{TRPCK},  &
\Rdef{TRUPCK}
\end{tabular}
\Usage
{\bf Choleski Decomposition} \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l}
{\tt CALL TRCHUL(S,W,M)} & $\mathbf{S = W'W}$ \\
{\tt CALL TRCHLU(S,V,M)} & $\mathbf{S = VV'}$
\end{tabular} \\[2mm]
{\tt S} is an $\mathtt{M \times M}$ {\it positive semi-definite}
symmetric matrix (e.g., error or weight matrix) and the routines
calculate the complementary lower triangular Choleski factors.
It is allowed to overwrite {\tt S} by {\tt W} or {\tt V}.
\par
{\bf Symmetric Multiplication of Lower Triangular Matrices} \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l@{ ${\bf \to}$ }l}
{\tt CALL TRSMUL(W,S,M)} & $\mathbf{W'W}$ & {\bf S} \\
{\tt CALL TRSMLU(W,R,M)} & $\mathbf{WW'}$ & {\bf R}
\end{tabular} \\[2mm]
{\tt W} is an $\mathtt{M \times M}$ lower triangular matrix and
{\tt S}, {\tt R} the two symmetric products of the multiplication of
{\tt W} by its transpose. It is allowed to overwrite {\tt W} by either
{\tt S} or {\tt R}.
\newpage
{\bf Lower Triangular Matrix Inversion} \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l@{ ${\bf \to}$ }l}
{\tt CALL TRINV(W,V,M)} & $\mathbf{W^{-1}}$ & {\bf V}
\end{tabular} \\[2mm]
{\tt W} is an $\mathtt{M \times M}$ lower triangular matrix which is
inverted into {\tt V} (the inverse of a lower triangular matrix is lower
triangular). {\tt W} {\it may have rows and columns of zeros}
as produced by the Choleski decomposition of a weight matrix with
unmeasured variables. It is allowed to overwrite {\tt W} by {\tt V}.
\par
{\bf Symmetric Matrix Inversion} \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l@{ ${\bf \to}$ }l}
{\tt CALL TRSINV(S,R,M)} & $\mathbf{S^{-1}}$ & {\bf R}
\end{tabular} \\[2mm]
{\tt S} is an $\mathtt{M \times M}$ {\it positive semi-definite}
symmetric matrix which is inverted into {\tt R} (also stored packed).
It is permissible to overwrite {\tt S} by {\tt R}.
\par
{\bf Triangular -- Rectangular Multiplication} \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l@{ ${\bf \to}$ }l}
{\tt CALL TRLA (W,A,B,M,N)} & $\mathbf{WA}$ & {\bf B} \\
{\tt CALL TRLTA(W,A,B,M,N)} & $\mathbf{W'A}$ & {\bf B} \\
{\tt CALL TRAL (A,V,B,M,N)} & {\bf AV} & {\bf B} \\
{\tt CALL TRALT(A,V,B,M,N)} & $\mathbf{AV'}$ & {\bf B}
\end{tabular} \\[2mm]
{\tt A} and {\tt B} are $\mathtt{M \times N}$ rectangular matrices,
{\tt W} is an $\mathtt{M \times M}$ lower triangular matrix, and
{\tt V} is an $\mathtt{N \times N}$ lower triangular matrix. In each
call it is allowed to overwrite {\tt A} by {\tt B}.
\par
{\bf Symmetric - Rectangular Multiplication} \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l@{ ${\bf \to}$ }l}
{\tt CALL TRSA (S,A,C,M,N)} & {\bf SA} & {\bf C} \\
{\tt CALL TRAS (A,R,C,M,N)} & {\bf AR} & {\bf C} \\
{\tt CALL TRSAT(S,B,C,M,N)} & $\mathbf{SB'}$ & {\bf C} \\
{\tt CALL TRATS(B,R,C,M,N)} & $\mathbf{B'R}$ & {\bf C}
\end{tabular} \\[2mm]
{\tt A} and {\tt C} are $\mathtt{M \times N}$ rectangular matrices,
{\tt B} is an $\mathtt{N \times M}$ matrix, {\tt S} is an
$\mathtt{M \times M}$ symmetrix matrix, and {\tt R} is an
$\mathtt{N \times N}$ symmetric matrix. It is {\it not} allowed to
overwrite {\tt A} or {\tt B} by the product matrix {\tt C}.
\par
{\bf Symmetric Multiplication of Rectangular Matrices} \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l@{ ${\bf \to}$ }l}
{\tt CALL TRAAT(A,S,M,N)} & $\mathbf{AA'}$ & {\bf S} \\
{\tt CALL TRATA(B,R,M,N)} & $\mathbf{B'B}$ & {\bf R}
\end{tabular} \\[2mm]
{\tt A} is an $\mathtt{M \times N}$ matrix, {\tt B} is an
$\mathtt{N \times M}$ matrix, {\tt S} is an $\mathtt{M \times M}$
symmetric matrix, and {\tt R} is an $\mathtt{M \times M}$ symmetric
matrix. No overwriting is allowed.
\par
{\bf Transformation of Symmetric Matrix } \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l@{ ${\bf \to}$ }l}
{\tt CALL TRASAT(A,S,R,M,N)} & $\mathbf{ASA'}$ & {\bf R} \\
{\tt CALL TRATSA(B,S,R,M,N)} & $\mathbf{B'SB}$ & {\bf R} \\
{\tt CALL TRQSQ (Q,T,R,M)}   & {\bf QTQ} & {\bf R}
\end{tabular} \\[2mm]
{\tt A} is an $\mathtt{M \times N}$ matrix, {\tt B} is an
$\mathtt{N \times M}$ matrix, {\tt S} is an $\mathtt{N \times N}$
symmetric matrix, and {\tt R}, {\tt Q}, {\tt T} are
$\mathtt{M \times M}$ symmetric matrices. No overwriting is allowed.
\par
{\bf Packing and Unpacking a Symmetric Matrix} \\[3mm]
\begin{tabular}{@{\hspace*{10mm}}l@{\qquad}l@{ ${\bf \to}$ }l}
{\tt CALL TRPCK (A,S,M)} & {\bf A} & {\bf S} \\
{\tt CALL TRUPCK(S,A,M)} & {\bf S} & {\bf A}
\end{tabular} \\[2mm]
{\tt A} is an $\mathtt{M \times M}$ unpacked symmetric matrix (all
$\mathtt{M}^2$ elements) and {\tt S} is the same matrix stored packed.
Overwriting is allowed for both {\tt TRPCK} and {\tt TRUPCK}.
\\ $\bullet$
