\Version{TRTCH}                  \Routid{M233}
\Keywords{CHARACTER DIFFERENT SET TRANSLATE CONVERSION UPPER LOWER}
\Keywords{EBCDIC ASCII}
\Author{A. Berglund, C. Osland}   \Library{KERNLIB, IBM and VAX only}
\Submitter{J. Richards}            \Submitted{01.12.1981}
\Language{Assembler}                \Revised{27.11.1984}
\Cernhead {Translate Between Different Character Sets}
{\tt TRTCH} consists of a series of general purpose routines which
will translate strings of any length according to the requested
translation table.
{\tt TRTCH} is mostly IBM only, however, the translations
between {\tt ASCII} and {\tt EBCDIC} are also available for VAX.
\Structure
{\tt SUBROUTINE} subprogram \\
User Entry Names: \Rdef{TRA0E0}, \Rdef{TRC0E0}, \Rdef{TRE0A0},
\Rdef{TRE0C0}, \Rdef{TRE0EL}, \Rdef{TRE0EU}, \Rdef{TRE0TN},
\Rdef{TRE0ZT}
\Usage
\begin{verbatim}
    CALL TRicoc(array,nlen)
\end{verbatim}
\begin{DLtt}{12345678}
\item [ic] Input character code (see table below).
\item [oc] Output character code (see table below).
\item [array] String of characters to be translated,
each character occupying one 8-bit byte.
\item [nlen] Number of characters to be translated.
\end{DLtt}
For character codes of $< 8$ bits, such as CDC display code,
right-justification with zero-fill is assumed. \\[3mm]
The character codes available are:
\begin{DLtt}{123456}
\item [A0] CERN standard {\tt ASCII},
\item [C0] CDC display code, as used at CERN,
\item [E0] CERN standard {\tt EBCDIC},
\item [EL] {\tt EBCDIC} with upper case letters converted to lower case,
\item [EU] {\tt EBCDIC} with lower case letters converted to upper case,
\item [TN] {\tt EBCDIC} with blanks for characters not on {\tt TN} train,
\item [ZT] {\tt EBCDIC} with dots, {\tt '.'}, for characters not on
{\tt TN} train.
\end{DLtt}
The available translation routines are:
\begin{DLtt}{12345678}
\item [TRA0E0] Standard {\tt ASCII} to standard {\tt EBCDIC}.
\item [TRC0E0] CDC display code to standard {\tt EBCDIC}.
\item [TRE0A0] Standard {\tt EBCDIC} to standard {\tt ASCII}.
\item [TRE0C0] Standard {\tt EBCDIC} to CDC display code.
\item [TRE0EL] Standard {\tt EBCDIC} to lower case only {\tt EBCDIC}.
\item [TRE0EU] Standard {\tt EBCDIC} to upper case only {\tt EBCDIC}.
\item [TRE0TN] Standard {\tt EBCDIC} to {\tt TN} train {\tt EBCDIC}.
\item [TRE0ZT] Standard {\tt EBCDIC} to {\tt TN} train {\tt EBCDIC}
with {\tt '.'}--fill.
\end{DLtt}
\Notes
The {\tt TN} train is the print train used by the IBM impact
printer. It contains upper and lower case characters and
a few unusual special characters such as boxes and numeric
superscripts.
\\ $\bullet$
