\Version{BLOW}                               \Routid{M426}
\Keywords{BLOW BYTE UNPACK FULL WORD}
\Author{CDC: J. Blake, G. Beltz, IBM: A. Berglund} \Library{KERNLIB}
\Submitter{}                              \Submitted{12.06.1972}
\Language{Fortran or Assembler} \Revised{01.02.1982}
\Cernhead {Unpack Full Words into Bytes}
{\tt BLOW} converts a source array containing a record consisting of a
continuous string of {\tt NBYTES} bytes of {\tt NBITS} bits per byte
into a target array of {\tt NBYTES} full words, right-adjusted with
zero-fill. {\tt BLOW} is the inverse of {\tt BUNCH} (M436).
\Structure
{\tt SUBROUTINE} subprogram \\
User Entry Names: \Rdef{BLOW} \\
External References: \Rind{UPKCH}{M427}
\Usage
\begin{verbatim}
    CALL BLOW(SOURCE,TARGET,NBYTES,NBITS)
\end{verbatim}
\begin{DLtt}{12345678}
\item [SOURCE] Source array containing the string of {\tt NBYTES} bytes.
\item [TARGET] Target array, which must be at least {\tt NBYTES} full
words long.
\item [NBYTES] Number of bytes in the source record ($\mathtt{0 < NBYTES}$).
\item [NBITS]  Number of bits per byte ($\mathtt{0 < NBITS \leq nbpw}$,
where $\mathtt{nbpw = 60}$ on CDC and $\mathtt{ = 32}$ on IBM).
\end{DLtt}
\Restrict
The two arrays {\tt SOURCE} and {\tt TARGET} must not overlap in any way.
\Errorh
{\tt BLOW} ignores calls with erroneous parameter values.
\Examples
CDC:
\begin{verbatim}
    CALL BLOW(SOURCE,TARGET,200,18)
\end{verbatim}
The array {\tt SOURCE} contains a record of 200 18-bit bytes, stored
contiguously in 60 60-bit words, i.e., a string of 3600 bits. After the
completion of the call to {\tt BLOW}, the array {\tt TARGET} will
contain 200 60-bit words, each containing one 18-bit byte,
right-justified with zero-fill.
\\ $\bullet$
