%  20 feb 1995  ksk
\Version{KERSET}                          \Routid{N001}
\Keywords{ERROR KERNLIB KERNNUM SECTION MONITOR}
\Author{H. Lipps}                            \Library{KERNLIB}
\Submitter{}                                  \Submitted{22.10.1984}
\Language{Fortran}                       \Revised{15.03.1993}
\Cernhead {Error Processing for Sections A-H of KERNLIB}
\begin{center}
\fbox{\parbox{120mm}{\POBSOLETE
Please note that, as a consequence of transferring subprograms
from KERNLIB to MATHLIB, this routine has been partially
obsoleted in CNL 211. It can, for a transitional period, still be
used for sections D ({\tt D509} only), and for sections E and
F of KERNLIB. Users are advised not to use it any longer for
other cases and to replace it in older
programs. With the foreseen transfer of the subroutines in sections
D,E,F in KERNLIB to MATHLIB, it will eventually disappear.
\\[3mm]
Suggested replacement: {\tt MTLSET} (N002)
}}
\end{center}
Subroutine {\tt KERSET} allows the user to redefine the action to be
taken by subprograms in the Fortran version of sections A-H of
KERNLIB when certain specified error conditions are detected.
(This subroutine does not exist in the Fortran 66 version.)
Error recovery may be performed either on each occurrence of the error,
or only a specified number of times. Messages may be written either on
each occurrence of the error, or only a specified number of times. Error
messages may be written (by default) onto the system output unit, or
may be re-routed to some other output file.
\Structure
{\tt SUBROUTINE} subprogram \\
User Entry Names: \Rdef{KERSET}\\
Internal Entry Names: {\tt KERMTR}\\
Files Referenced: Printer or user-defined\\
External References:  \Rind {ABEND}{Z035}
\Usage
\begin{verbatim}
    CALL KERSET(ER,LGFILE,LM,LR)
\end{verbatim}
\begin{DLtt}{12345678}
\item [ER] ({\tt CHARACTER*6}) A character string that identifies the
range of error conditions for which action is to be redefined.
\item [LGFILE]({\tt INTEGER}) The logical unit number to be used
for error messages, or zero if error messages are to be written
onto the system output unit.
\item [LM] ({\tt INTEGER}) The number of occurrences of each error
condition in the range {\tt ER} for which an error message is to be
written. $\mathtt{LM \leq 0}$ is treated as zero, $\mathtt{LM \geq 100}$
as infinity.
\item [LR] ({\tt INTEGER}) The number of times that error
recovery is to be performed for each error condition in the range
{\tt ER}. $\mathtt{LR \leq 0} $ is treated as zero.
$\mathtt{LR \geq 100}$ is treated as infinity. If any error condition in
the range {\tt ER} occurs $\mathtt{LR+1}$ times a message is printed and
the run is terminated by calling {\tt ABEND} (Z035).
\end{DLtt}
\newpage
\Notes
\begin{enumerate}
\item {\tt KERSET} applies to those KERNLIB error conditions which are
specified by a six-character code (e.g., {\tt C204.2}) in the
{\bf Error handling} section of the Short Write-ups.
\item If the string {\tt ER} consists of six characters specifying a
single error condition \\
(e.g., {\tt ER}$=${\tt 'C204.2'}), {\tt LM} and
{\tt LR} apply only to this one error condition.
\par
If the six-character string {\tt ER} ends with one or more blanks,
{\tt LM} and {\tt LR} apply to all error conditions whose leftmost
characters match the non-blank characters of {\tt ER}. \\
Thus {\tt ER $=$ 'C2\ \ \ \ '}  (four blanks) applies to all error
conditions in packages C200 to C299, and {\tt ER $=$ '\ \ \ \ \ \ '}
(six blanks) applies to all error conditions under the control of
{\tt KERSET}.
\item The value of {\tt LGFILE} applies to all error messages written
under the control of {\tt KERSET}.
\end{enumerate}
$\bullet$
