% 27 feb 95  ksk
\Version{RDJMNB}               \Routid{U501}
\Keywords{BETA-TERM B-TERM WIGNER D-FUNCTION FINITE ROTATION OPERATOR}
\Author{K.S. K\"olbig}           \Library{MATHLIB}
\Submitter{}                    \Submitted{15.02.1989}
\Language{Fortran}        \Revised{01.12.1994}
\Cernhead{Beta-Term in Wigner's D-Function}
Function subprograms {\tt RDJMNB} and {\tt DDJMNB} calculate
the $\beta$-term $d _{mn}^j(\beta)$ in the matrix element of the finite
rotation operator (Wigner's $D$-function)
$$ D_{mn}^j(\alpha,\beta,\gamma) \ = \
\displaystyle e^{-im\alpha} \ d_{mn}^j(\beta) \ e^{in\gamma} $$
by using the formula (Ref. 1, No. 4.3.1(3)) \\ [4mm]
\hspace* {35mm}
$ \displaystyle d_{mn}^j(\beta) \ = \ \displaystyle
(-1)^{j+m} \sqrt{(j+m)!(j-m)!(j+n)!(j-n)!} \ \times $ \\ [4mm]
\hspace* {55mm}
$ \displaystyle \sum_k \ (-1)^k \
\frac{\cos^{2k-m-n}(\textstyle \frac{1}{2}\beta) \
\sin^{2j+m+n-2k}(\textstyle \frac{1}{2}\beta)}
{k!(j+m-k)!(j+n-k)!(k-m-n)!} $ \\ [4mm]
for arbitrary (either all integer or all half-integer) values of $j,m,n$
such that $j \ge 0, |m| \le j$ and $|n| \le j$. The summation over $k$
runs from $\max(0,m+n)$ to $\min(j+m,j+n)$.
\par
On computers other than CDC or Cray, only
the double-precision version {\tt DDJMNB} is available.
On CDC and  Cray computers, only the single-precision version
{\tt RDJMNB} is available.
\Structure
{\tt FUNCTION} subprograms \\
User Entry Names: \Rdef{RDJMNB}, \Rdef{DDJMNB}\\
Obsolete User Entry Names: \Rdef{DJMNB} $\equiv$ \Rdef{RDJMNB} \\
Files Referenced: {\tt Unit 6} \\
External References: \Rind{MTLMTR}{N002}, \Rind{ABEND}{Z035}
\Usage
In any arithmetic expression,
\begin{center}
{\tt RDJMNB(AJ,AM,AN,BETA)} \quad or \quad {\tt DDJMNB(AJ,AM,AN,BETA)}
\quad has the value \quad $d_{mn}^j(\beta)$,
\end{center}
where $\mathtt{AJ} = j$, $\mathtt{AM} = m$, $\mathtt{AN} = n$ and
$\mathtt{BETA} = \beta$.
{\tt RDJMNB} is of type {\tt REAL}, {\tt DDJMNB} is of type
{\tt DOUBLE PRECISION}, and {\tt AJ}, {\tt AM}, {\tt AN}, {\tt BETA}
have the same type as the function name. {\tt BETA} has to be given in
degrees.
\Restrict
$\mathtt{0 \le AJ \le 25}$, $\mathtt{|AM| \le AJ}$,
$\mathtt{|AN| \le AJ}$, $\mathtt{0 \le BETA \le 360}$.
\Accuracy
Approximately full single- or double-precision machine accuracy, at
least for small values of the indices.
\Errorh
Error {\tt U501.1}: If any of the restrictions is not satisfied,
the function value is set equal to zero, and a message is written on
{\tt Unit 6}, unless subroutine {\tt MTLSET} (N002) has been called.
\Refer
\begin{enumerate}
\item D.A. Varshalovich, A.N. Moskalev and V.K. Khersonskii,
Quantum theory of angular momentum, (World Scientific, Singapore 1988) 76
\end{enumerate}
$\bullet$
