\Version {RNDM2}                       \Routid{V107}
\Keywords{DISTRIBUTION GENERATOR IBM RANDOM NUMBER UNIFORM}
\Author{McGill University}             \Library{MATHLIB, IBM and VAX only}
\Submitter{T. Lindel\"of}             \Submitted{15.09.1978}
\Language{CDC: Fortran, IBM: Assembler}  %\Revised{}
\Cernhead {IBM Uniform Random Number Generator}
{\tt RNDM2} is a combined multiplicative congruential generator and
a shift register generator with nearly optimal lattice structure in
2, 3, 4 and 5 dimensions, with a period of about $5 \times 10^{18}$
numbers, designed especially for the IBM where the 32-bit word length
makes the usual generators ({\tt RNDM}, {\tt RN32}) not good enough for
some calculations.
A Fortran version of {\tt RNDM2} exists on CDC for compatibility
reasons, but the CDC version is not recommended for production
use. It only functions as an interface to {\tt RNDM} (V104).
\Structure
{\tt FUNCTION} subprogram \\
User Entry Names: \Rdef{RNDM2}, \Rdef{IRNDM2}, \Rdef{RD2IN},
\Rdef{RD2OUT}
\Usage
\begin{verbatim}
    U = RNDM2(DUMMY)
\end{verbatim}
assigns to {\tt U} a random normalized floating point value in the
interval (0,1), with uniform distribution.
\begin{verbatim}
    K = IRNDM2(DUMMY)
\end{verbatim}
assigns to {\tt K} a random integer value in the interval $(0,2^{31})$,
with uniform distribution.
\begin{verbatim}
    CALL RD2OUT(IS,JS)
\end{verbatim}
sets {\tt IS} and {\tt JS} to the current values of two internal
registers used in the congruential generator and the shift register
generator respectively.
\begin{verbatim}
    CALL RD2IN(IS,JS)
\end{verbatim}
resets the two above-mentioned registers to {\tt IS} and {\tt JS}
respectively. The entire future sequence of generated numbers is a
function of {\tt IS} and {\tt JS}.
\par
In order to avoid loss of randomness it is advisable to use for
{\tt IS} and {\tt JS} only values obtained through an earlier call to
{\tt RD2OUT}. Either parameter may be set to zero in order to reduce the
generator to a pure shift register generator ($\mathtt{IS=0}$) or to a
pure congruential generator ($\mathtt{JS=0}$).
$\mathtt{IS=JS=0}$ produces a sequence of only zeros.
\Refer
\begin{enumerate}
\item G. Marsaglia, K. Ananthanarayanan and N. Paul,
Comm. ACM {\bf l5} (1972) 873.
\end{enumerate}
$\bullet$
