\Version {GENBOD}                       \Routid{W515}
\Keywords{MANY-BODY EVENT MONTE-CARLO GENERATOR}
\Author{F. James}                        \Library{POOL}
\Submitter{}                             \Submitted{20.10.1975}
\Language{Fortran}                    %\Revised{}
\Cernhead {N-Body Monte-Carlo Event Generator}
{\tt GENBOD} generates a multi-particle weighted event according to
Lorentz-invariant Fermi phase space. It is a modification of the
routine {\tt GENEV} (in {\tt FOWL} (W505)) and uses the method of
Raubold and Lynch (see Ref. 1).
The total CM energy as well as the number and masses of the
outgoing particles are specified by the user, but may be changed
from event to event. {\tt GENBOD} generates the CM vector momenta
(and energies) of the outgoing particles and gives the weight which
must be associated with each event. The weight may then be
multiplied by any 'matrix element' or geometrical detection function
calculated by the user.
\Structure
{\tt SUBROUTINE} subprogram \\
User Entry Names: \Rdef{GENBOD}\\
Files Referenced: Printer\\
External References: \Rind{FLPSOR}{M103}, \Rind{RNDM}{V104},
\Rind{PDK}{W505}, \Rind{ROTES2}{W505} \\
{\tt COMMON} Block Names and Lengths : {\tt /GENIN/ 21},
{\tt /GENOUT/ 91}
\Usage
\begin{verbatim}
    COMMON /GENIN /NP,TECM,AMASS(18),KGENEV
    COMMON /GENOUT/ PCM(5,18),WT
    CALL GENBOD
\end{verbatim}
{\bf Input:}
\begin{DLtt}{12345678}
\item [NP] ({\tt INTEGER}) Number of outgoing particles
$\mathtt{(2 \leq NP \leq 18)}$.
\item [TECM] ({\tt REAL}) Total CM energy.
\item [AMASS]({\tt REAL}) Array where element {\tt I} contains the
mass of the {\tt I}-th outgoing particle.
\item [KGENEV] ({\tt INTEGER})
$\mathtt{= 1}$ for cross section constant with energy,
$\mathtt{= 2}$ for Fermi energy dependence.
\end{DLtt}
{\bf Output:}
\begin{DLtt}{1234567890}
\item [PCM(1,I)] ({\tt REAL}) $P_x$ of {\tt I}-th particle.
\item [PCM(2,I)] ({\tt REAL}) $P_y$ of {\tt I}-th particle.
\item [PCM(3,I)] ({\tt REAL}) $P_z$ of {\tt I}-th particle.
\item [PCM(4,I)] ({\tt REAL}) Energy of {\tt I}-th particle.
\item [PCM(5,I)] ({\tt REAL}) $P$ of {\tt I}-th particle.
\item [WT] ({\tt REAL}) Weight of the event.
\end{DLtt}
See also the {\bf Long Write-up} for {\tt FOWL} (W505).
\Refer
\begin{enumerate}
\item F. James, Monte Carlo Phase Space, CERN 68-15 (1968)
\end{enumerate}
$\bullet$
