\Version{IUCHAN}                     \Routid{Y201}
\Keywords{CHANNEL FIND HISTOGRAM}
\Author{J. Zoll, P. Rastl}                      \Library{KERNLIB}
\Submitter{C. Letertre}               \Submitted{01.09.1969}
\Language{Fortran or Assembler}       \Revised{16.09.1991}
\Cernhead {Find Histogram-Channel}
{\tt IUCHAN}, {\tt IUBIN}, {\tt IUHIST} all find the histogram-channel
for a given quantity in the same way.
They differ only slightly in the way in which the parameters are passed.
\Structure
{\tt FUNCTION} subprograms   \\
User Entry Names: \Rdef{IUCHAN}, \Rdef{IUBIN}, \Rdef{IUHIST}
\Usage
All routines need the the following parameters:
\begin{DLtt}{123456}
\item [X] ({\tt REAL}) Quantity to be histogrammed.
\item [XLOW] ({\tt REAL}) Lower limit of the histogram.
\item [DX] ({\tt REAL}) Channel width.
\item [NX] ({\tt INTEGER}) Number of channels.
\end{DLtt}
and they return the channel number
$\mathtt{N=(X-XLOW)/DX+1+\varepsilon}$ normally,
or $\mathtt{N=0}$ for underflow ($\mathtt{X<XLOW}$),
or $\mathtt{N=NX+1}$  for overflow ($\mathtt{X \geq XLOW + NX*DX}$).
\par
$\varepsilon > 0$ is a small bias to counteract rounding effects when
{\tt X} is exactly on a bin edge, a likely and serious problem when
compressed data are histogrammed.
\par
$\varepsilon = 10^{-5}$ on 32-bit machines,
$\varepsilon = 10^{-6}$ on machines with a larger word size. \\[3mm]
{\bf Function IUCHAN:}
\begin{verbatim}
    N = IUCHAN(X,XLOW,DX,NX)
\end{verbatim}
\par
{\bf Functions IUBIN and IUHIST:}
\begin{verbatim}
    DIMENSION PAR(3)
    EQUIVALENCE(NX,PAR(1))
    LOGICAL SPILL
 
    N = IUBIN (X,PAR,SPILL)
    N = IUHIST(X,PAR,SPILL)
\end{verbatim}
with
\begin{DLtt}{123456}
\item [PAR] Histogram parameters: \\
$\mathtt{PAR(1) \equiv NX}$  \\
$\mathtt{PAR(2) \equiv DX}$ (for {\tt IUBIN}),
or $\mathtt{\equiv 1/DX}$ reciprocal of the channel width (for
{\tt IUHIST}). \\
$\mathtt{PAR(3) \equiv XLOW}$
\item [SPILL] ({\tt LOGICAL}) Flag set to {\tt .TRUE.} or {\tt .FALSE.}
depending on whether {\tt X} is outside
or inside the histogram.
\end{DLtt}
$\bullet$
