%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                 %
%  GEANT manual in LaTeX form                                     %
%                                                                 %
%  Michel Goossens (for translation into LaTeX)                   %
%  Version 1.00                                                   %
%  Last Mod. Jan 24 1991  1300   MG + IB                          %
%                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Documentation{S.Giani, S.Ravndal}
\Submitted{22.03.94}      \Revised{22.03.94}
\Version{Geant 3.21}\Routid{GEOM020}
\Makehead{``MANY'' Volumes and boolean operations on volumes}

The topology of 
{\tt GEANT} is given by 16 general shapes which are used to represent
solids in the design of detectors. These shapes are bounded by 2-nd order
surfaces for reasons of speed at tracking time. New tools have been 
introduced to extend the power of the {\tt GEANT} geometry: boolean operations 
(union, intersection and subtraction) between such shapes are now allowed, so
that an infinity of new shapes can be built; divisions along arbitrary axis
are now possible, allowing to exploit any simmetry in the design.
The tracking is now able to handle overlapping volumes in the most general
sense: the concept of {\tt MANY} volumes has been extended to automatic 
clipping 
for protuding objects. Thanks to this {\tt MANY} technique, 
any 3D or 2D geometrical
pattern can be reduced to a single dimensional structure, basically cancelling
the need for {\tt GUNEAR} at tracking time. 
Finally, more than one order of magnitude in speed has been gained compared
to the old step search algorithm used to track in {\tt MANY} volumes. \\[.5cm]
\begin{itemize}
\item Topological extensions via boolean operations.
\item Divisions along arbitrary axis.
\item Reduction of 2D structure to 1D structure.
\item Logic and algorithm.
\end{itemize} 
 
\section{Topological extensions via boolean operations}

Boolean operations between a given set of shapes allow the creation of an
infinite number of shapes. Adding a single new primitive (for ex. a toroid) to 
the basic set of shapes, all the possible combinations with all the old shapes
become moreover possible.\\[.5cm]
\begin{itemize}
\item UNION: the union of two volumes B and C is obtained positioning two
             overlapping `{\tt MANY}' daughters B and C inside the same mother
             A.
             To identify the result of the union as a single volume is enough
             to create a SET associated to the volumes B and C. 
\item SUBTRACTION: the subtraction of a volume B from a volume C is obtained
                   positioning B as `{\tt ONLY}' and C as a `{\tt MANY}' 
                   overlapping B.
                   The result of the subtraction is automatically what the
                   tracking sees as the C volume, so {\tt GSPOS} is the only 
                   user
                   interface needed. The volume B can have the same material as
                   the mother A or not. The normal positioning technique used
                   for `{\tt ONLY}s' is a particular case of boolean subtraction (a
                   volume is contained inside the other).
\item INTERSECTION: the intersection of two volumes B and C is obtained 
                    positioning a protuding `{\tt MANY}' object C in 
                    a `{\tt MANY}' object
                    B which is normally placed in a `{\tt ONLY}' mother A. A and B
                    must have the same material, while C has the material asked
                    for the intersection. The intersection is given by the part
                    of C which is not protuding from B, therefore by what the
                    new tracking sees as the volume C: again the only needed
                    user interface is {\tt GSPOS}. If not interacting with other
                    daughters of A, B could also be `{\tt ONLY}'. Without any
                    ambiguity, even further `{\tt MANY}' daughters of A can overlap
                    the protuding part of C outside B, because it is really
                    invisible to the new tracking. 
\end{itemize}
 
\section{Divisions along arbitrary axis}

As protuding {\tt MANY} daughters are automatically clipped by their mother, 
it is
possible to divide objects along arbitrary directions. Suppose for example
that we want to divide a PCON into planar slices parallel to the plane
X + Y = 0; it is enough to position a {\tt MANY} BOX embedding completely the
PCON `inside' the PCON itself; the BOX must be rotated of 45 degrees in X and
Y and then divided along one of its natural axis. Another example can be the
following: suppose we want to divide 24 times in PHI a PGON with only 6 natural
segmentations in PHI; 
then we have to position a {\tt MANY} TUBE, completely embedding
the PGON, `inside' the PGON itself; again, we can now divide the tube in 24
slices in PHI and the tracking will see only the PGON divided 24 times in PHI.
\begin{center}
Reduction of 2D structure to 1D structure.
\end{center}
Sometimes there are complicated multidimensional geometrical structures (like
honeycombs, spaghetti, etc.) which can be reduced to the overlap of several
one-dimensional patterns (for which divisions can be used or virtual divisions
are particularly efficient). The {\tt MANY} option allows to superimpose such 1D
structures to reproduce the multi-dimensional ones in a very effective way,
instead of positioning the volumes one by one in the 2D or 3D pattern.
The following picture will explain better this concept.


\section{Logic and algorithm}

In {\tt GTMEDI}, all the '{\tt MANY}' volumes for which the point 
is found inside (and not
in any of its daughters) is put in a stack saving also its branch of the tree.
The point is considered to be in the deepest level 
'{\tt MANY}'. Each time the point
is also found in an intermediate '{\tt ONLY}' volume the stack is 
cleared (to insure
boolean subtraction). In {\tt GTNEXT}, if stepping 
inside a '{\tt MANY}' volume, not {\tt ONLY}
its daughters and its own boundaries are checked, but also the daughters and
the boundaries of all the others '{\tt MANY}' in the 
list have to be checked. This
is not yet enough: the logic requires that the boundaries of all the possible
overlapping volumes are checked, so it is also necessary for each `{\tt MANY}' 
volume 
of the list to go up to the first '{\tt ONLY}' mother 
in the tree repeating the same 
procedure. This allows also boolean intersection. Obviously, volumes already
checked are not checked again. 


