\chapter{CONDITIONS OF AVAILABILITY}
Program Availability and Charging
\par The \ttbf{HBOOK} package has been installed on the following computers:
IBM3090, VAX, NORD$>-$500, UNIVAC, APOLLO, CRAY, CYBER205.
\par \ttbf{HBOOK} is part of the CERN Program Library and may be used freely inside CERN.
Institutes collaborating with CERN and Physics departments in
member state universities or in National Research labs
in Member States may receive \ttbf{HBOOK} and its
documentation freely and without charge, subject only to a limitation
on the number of copies to be sent to any one institution free of charge.
\par CERN reserves the right to charge a higher fee, up to the full
cost of developing the programs, or to refuse any request
which it deems to be not in the interest of the Organization.
\begin{center}{\bf Conditions of Use}\end{center}
\par Programs and documentation are provided solely for the use of
the organization to which they are distributed and may not be redistributed
to any third party without the express agreement of CERN.
\par The material cannot be sold.
\par CERN should be given credit in all references and publications
based on the programs.
\par CERN undertakes no obligation for maintenance of the programs,
nor responsibility for their correctness,
and accepts no liability whatsoever resulting from the use of its programs.
Although we do not ''support'' the programs in a commercial sense,
we will answer questions concerning the implementation or usage
of the programs and we welcome suggestions for improvements.
\par Requests for \ttbf{HBOOK} and full details of its availability
should be addressed to:
\begin{verbatim}
                                         The Program Librarian
                                         Data Handling Division
                                         CERN
                                         CH 1211 GENEVA 23
\end{verbatim}
\par At CERN the \ttbf{HBOOK} library is part of the
\ttbf{PACKLIB} library.
\index{PACKLIB}
 
