%.*****************************************************************>
%.* VAXTAP Reference Guide - Overview                             *>
%.*  Last Mod.   24 Jan 1992  10:40   mg                          *>
%.*****************************************************************>
\Filename{H1vaxover-Introduction}
\chapter{Introduction}
\par
The utilities described in this manual were originally developed
for the CERN VAXcluster VXCERN.

Several essential features were felt to be missing from standard
VAX tape handling utlities. The most important of these were

\begin{UL}
\item
Support for IBM SL tape labels
\item
Support for visual identifiers (VIDs)
\item
Efficient device allocation
\end{UL}

To this end, a new utility was written, known as SETUP. This
command, which uses standard VAX/VMS system services, performs
device allocation, tape mounting and label checking.

This command was soon joined by the other utilities described in
this manual, namely LABELDUMP, STAGE, TAPECOPY, WRTAPE and XTAPE.

An interface to the CERN Tape Management System exists, and is
activated at installation time as described in the installation
chapter of this manual.

In addition, and interface from the FATMEN File and Tape Management
system is also provided. Further details can be found in the FATMEN
manual~\cite{bib-FATMEN}.

Originally, the commands described in this manual were defined
using the DCL {\tt SET COMMAND} utility. However, they have
recently been rewritten so that this is no longer required,
using the techniques described in~\cite{bib-NOCLD}.

