\chapter{Introduction}
 
    This manual describes the CERN implementation of the
Numerical Algorithms Group Program Library. This includes the NAG
Fortran Library, the NAG Graphics Library, the Online Information
Supplement, the PC/Workstation Library and the PC Graphics Library. More
detailed information can be found in Appendix A.
 
This document is provided solely for internal use at CERN, and should be
sufficient for most users of the library. Main parts of the introductory
guide to the mathematical part of the library are reproduced in Appendix
B , by kind permission of NAG Ltd. Reference copies of the complete
manuals are available in the Computer Science Library
and
the Theory Division Library for inspection. Additional sets have been
distributed to several groups. Type
\Ucom{XFIND NAG MANUALS}
on CERNVM or VXCERN for the precise location of all sets.
Users interested in buying their own copies are referred to
contact NAG directly:
 
\begin{verbatim}
     NAG Response Center
     Telephone: +44 865 311744
     Telefax: +44 865 311755
\end{verbatim}
 
  The NAG organization has granted CERN a site-wide licence for all
systems, for which there is an application available, as they are listed
in Appendix A. Users wishing to install the NAG library on their local
machine should contact the CERN Program Library Office.
 
  An interactive help facility called NAGHELP, which includes
access to the subroutine calling sequence
descriptions (for both the graphical and the mathematical parts of the
NAG library) is also provided by NAG. This facility has been installed
on CERNVM, VXCERN, CERNAPO and CSF services.
In addition, the same documentation is available on VXCERN in the
usual VMS style.
 
The way to access the online help documents on the
different CERN services is described below. We recommend the use of
NAGHELP on those systems where it is implemented, since it provides
access to the full description of the routines.
\begin{center}
\begin{tabular}{|*{3}{c|}}
\hline
\bf SYSTEM & \bf NAG Tool & \bf Further Information \\
\hline
\bf CERNVM &\Ucom{NAGHELP} &\Ucom{XFIND NAG}\\[1mm]
\hline
\bf VXCERN &\Ucom{NAGHELP} &\Ucom{HELP NAG} or \Ucom{XFIND NAG}\\
    &\Ucom{HELP @NAGLIB} &and the .DOC files in the NAG directories\\
\hline
\bf UNIX &\Ucom{NAGHELP} & files in the NAG directory\\
\hline
\end{tabular}\\
\end{center}
 
 
  There are 932 examples on the NAG directories which show the use
of the Fortran Library and of the Graphics Library. In the case of the
mathematical part, the expected results of the given tests are also
available for comparison. For the graphics examples, the 74 example
results are shown in Appendix D. How to access and run these examples is
described in Chapters 1.3 and 2.3.
 
