\chapter{Intended audience}

This manual is aimed primarily at those people responsible for the
installation of the CERN Program Library on systems at CERN or
elsewhere. Some of the material is only relevent for the
CERN Program Library section of the CN/ASD group at CERN.

Having read this manual you should be able to:

\begin{enumerate}
\item
Install all or part of the CERN Program Library on one of the 
supported platforms.
\item
Understand how to port the Program Library to a new platform.
\end{enumerate}

Certain sections of this manual will also be of interest
to software developers, such as Appendix \ref{sect-FLAGS},
where the C pre-processor flags are described and Appendix \ref{sect-PREFIX},
where reserved routine and common block prefixes are listed.

\chapter{Overview of the CERN Program Library and related environment}
The CERN Program  Library is a large collection  of general purpose programs
maintained  and offered  in both  source and  object code  form on  the CERN
central computers.   Most of these programs  were developed at CERN  and are
therefore  oriented towards  the  needs of  a  physics research  laboratory.
Nearly all, however, are of  a general mathematical or data-handling nature,
applicable to a wide range of problems.
 
\section{Contents and Organization of the Library}
The library contains several thousand subroutines and complete programs which are
grouped  together  by  logical  affiliation into several hundred program
packages. 80\%  of the programs are  written in FORTRAN and  the remainder in
assembly code, or C  usually with  a FORTRAN  version also  available.  The
language supported is currently Fortran 77.
 
Each package is assigned a unique code, consisting of a letter
and three or four digits. The  letter is used to indicate the category
in which the program or package resides. A package  consists of one or more related
subprograms with one package name and one or more user-callable entry names,
all described briefly in a {\tt Short write-up}~\cite{bib-CERNLIB}, and if necessary, an additional
{\tt Long  write-up}.  The terms under which the library material and associated
documentation may be distributed are given below.
 
\section{Availability and Charging}
\begin{itemize}
\item
Access to the CERN Program Library is free of charge to all HEP users
worldwide.
\item
Network  access  is the  sole distribution method.  
In order to gain  access, users must register the node
name of  their computer with  the Program  Library Office.  This  service is
free of charge.  Distribution on magnetic tape is not available.
\item
Non HEP academic and not-for-profit organizations are offered
registered network access for one  year for a registration fee of 1000
Swiss Francs. This service is free for physics departments of
institutes in CERN member states.
\item
Commercial users are offered 
registered  network access  for 1 year  for a base  fee of  5000 Swiss Francs.

The above fees are doubled for requests coming from non-member states.
\end{itemize}
 
The  Library  is  not  available  for  any  purpose   related  to  military
applications.
 
Distribution  to  organizations  not  covered  by the  above  rests  at  the
discretion  of  the  Director-General.    Machine  readable  copies  of  the
documentation are provided with the library and recipients are free to make
copies  for  their  users.   Paper  copies of  the  documentation  are  made
available  for  a  charge  which  covers the  printing  and  handling  cost.
Charging is waived for cumulative orders of less than 100 Swiss Francs.
 
If  your organisation is  expected to pay a  fee as
described above,  it will be  billed by CERN after the material
has been shipped. Please include special
billing address and instructions, if any, with the request.
 
As of the  publication date of this  document the member states  of CERN are
Austria, Belgium, Czech Republic, Denmark, Finland, France, Germany, Greece,
Hungary, Italy, The Netherlands, Norway, Poland, Portugal, Slovak Republic, Spain, Sweden,
Switzerland and the United Kingdom.  

States with Observer status are Israel, the Russian Federation, Turkey, Yugoslavia (status
suspended after UN embargo, June 1992), the European Commission and Unesco.  
 
\section{Conditions of Use Outside CERN}
 
Programs  and  documentation are  provided  solely  for  the  use of  the 
organization to which  they are distributed, and may not  be redistributed or
reproduced in  large numbers without the  express agreement of  CERN.  Note
that such agreement may have to be established somewhere else in addition to
or instead of CERN in the case of programs originating from sources outside
CERN.  The  appropriate Short  Write-up gives information  about authorship.
The material cannot be sold.  CERN should be given credit in all references,
library documentation, and publications based on the programs.
 
If the programs are  modified beyond what is necessary to  adapt them to the
local  machine/system  environment,  it  should   be  made  clear  in  local
documentation that they are locally modified versions of the CERN originals.
CERN should be informed of such modifications, and given the possibility of
introducing  the  same modifications  in  the  original version.   If  local
modifications are so  important as to change significantly  the behaviour of
the program, its  name should also be  changed in order to  avoid confusion
with the  original.  CERN welcomes  comments concerning the  Program Library
service, but undertakes  no obligation for the maintenance  of the programs,
nor  responsibility   for  their  correctness,  and   accepts  no  liability
whatsoever resulting from the use of its programs.
 
\chapter{CERNLIB installation environment}

{\bf N.B. from release 96A of the CERN Program Library, the installation
procedures have been completely rewritten to

\begin{itemize}
\item
Use standard tools, rather than HEP specific packages
\item
Permit common procedures across multiple systems (Unix, Windows/NT, VMS)
\end{itemize}

HEP specific packages such as PATCHY and CMZ are no longer used.

}

In most cases, such as on the CERNSP system at CERN, the manual
page for any of the following components can be obtained by
typing, for example, \underline{man cvs}.

\section{CVS}

\htmladdnormallinkfoot{CVS}{http://asdwww.cern.ch/pl/cvs/index.html}, or the
Concurrent Versions System, is a system that lets groups of people work 
simultaneously on groups of files (for instance program sources). 

It works by holding a central `repository' of the most recent version of the files. You may
at any time create a personal copy of these files by `checking out' the files from the
repository into one of your directories. If at a later date newer versions of the files are put
in the repository, you can `update' your copy. 

You may edit your copy of the files freely. If new versions of the files have been put in the
repository in the meantime, doing an update merges the changes in the central copy into
your copy. 

When you are satisfied with the changes you have made in your copy of the files, you can
`commit' them into the central repository. 

When you are finally done with your personal copy of the files, you can `release' them and
then remove them. 

From release 96A of the CERN Program Library, the master sources of the library 
are maintained in CVS. However, you do not need access to CVS to build the
library yourself - this is done using 'checked -out' files - pure Fortran and/or C
files, which include CPP statements to handle conditional compilation (e.g. machine
dependancies) and include files.

\section{Imake}

Imake is a tool that was written by the X consortium and is used to build
the X software. 

Imake is used to generate Makefiles from a template, a set
of  cpp  macro  functions,  and a per-directory input file
called an Imakefile.   This  allows  machine  dependencies
(such  has  compiler options, alternate command names, and
special make rules) to be kept separate from the  descriptions of the various items to be built.

\section{Makedepend}

Makedepend was also written by the X consortium is used to
determine the dependencies that a given soruce file
has on various include files, either directly or indirectly.

Makedepend is normally invoked by having a "depend" target
in the makefile to be updated - typing \underline{make depend}
brings the dependencies in the makefile up to date.

\section{Make}

Make is a tool that allows you to rebuild programs
from their components. The {\it rules} for building {\it targets}
are described in a {\it makefile}. The make program ensures
that only those steps that have to be repeated are executed.
Thus, if none of the source files that make up a module have changed, 
they are not recompiled. 

\section{CPP}

The C preprocessor, or CPP, permits conditional compilation and 
performs file inclusion and macro substitution on
source files. It is automatically invoked by the C compiler, and also by
the Fortran compiler on many Unix systems (normally requiring an extension
.F, rather than .f).

\chapter{Space requirements}

\index{disk space}

The complete CERN library requires approximately 200MB of disk
space. Slightly over 50MB may be saved if the sources are not
kept locally.

\chapter{Description of CERNLIB components}

The CERN Program Library consists of a number of independent
Fortran callable libraries and a collection of complete
programs. The overall structure is described briefly below.
Note that the library is not static and small deviations
from what is described may exist.

\section{KERNLIB}

\KERNLIB{} is a Fortran callable library composed almost entirely
of individual subroutines and functions. All of these are described
in the CERNLIB short writeups manual~\cite{bib-CERNLIB}.

The \KERNLIB source files are organised as follows:
\begin{itemize}
\item
Machine independent routines, mainly Fortran.
\item
Machine dependent routines, often C or Assembler.
\item
Numerical routines, including random number generators.
\end{itemize}

The actual contents of the library varies slightly from
platform to platform, largely due to the machine 
dependent routines.

%A complete list of routines, grouped according to section,
%is given below.
%
%\include{/afs/cern.ch/project/cnas_doc/sources/shortwrups/crnlbcat.tex}

\section{MATHLIB}

\MATHLIB{} is a Fortran callable library containing routines
of a mathematical nature. It is made up of four components,
namely

\begin{itemize}
\item
BVSL - basic vector subroutine library. These routines
were originally written for the IBM 3090 vector facility.
However, a Fortran version is now available.
\item
GEN - general mathematical routines.
\item
LAPACK - the well-known linear algebra package.
\item
MPA - multiple precision floating point arithmetic routines.
\end{itemize}

Note that LAPACK and MPA may only be distributed in object form
by CERN. Further details can be found in section \ref{sect-LAPACK}
and \ref{sect-MPA}.

\section{PACKLIB}

\PACKLIB{} is a Fortran callable library made up of complete
packages. A few of the packages, such as VMIO, are
machine dependent. The bulk, however, run on a variety
of systems including Unix, MS-DOS, Windows NT, VMS, VM/CMS
and MVS.

Long writeups exist for each of these packages and these
should be consulted for further details.

% Link to appropriate long writeup?

\begin{itemize}
\item
EPIO - EP standard input/output package
\item
FFREAD - processing of Free Format data cards
\item
IOPACK - Input/output package for VM/CMS and MVS systems
\item
KAPACK - Key access package
\item
KUIP - Kit for a User interface package
\item
HBOOK - Histogramming, statistical analysis and Ntuple package
\item
MINUIT - Function minimization package
\item
VMIO - VM/CMS specific I/O package
\item
ZBOOK - Data structure management package
\item
ZEBRA - Data structure management package
\item
CDLIB - The API for the HEPDB detector geometry and calibration system
\item
FATLIB - The API for the FATMEN distributed file management system
\end{itemize}


\section{The graphics libraries}

The graphics libraries are divided into a kernel and
driver specific libraries. Examples are shown below.

\begin{itemize}
\item
GRAFLIB - the graphics kernel
\item
GRAFGKS - the GKS binding
\item
GRAFGDDM - the GDDM binding
\item
GRAFX11 - the X11 binding
\end{itemize}

Some of the above libraries may not be available
on certain platforms.

\section{PAWLIB}

\PAWLIB{} is the library used to build the various
\PAW{} modules. In addition to \PAW{} itself, (apart
from the main program), it also contains
\COMIS{} and \SIGMA{}.

\section{GEANT}

\GEANT{} is provided in a standalone library. The version
number is contained in the library name.

\section{The Monte Carlo libraries}

The Monte Carlo libraries provide numerous event
generators, such as the well-know Lund Monte Carlos,
including those listed below. Again, the long writeup
should be consulted for further details.

% Link?

\begin{itemize}
\item
COJETS - generator for high energy proton-proton and
proton-antiproton collisions
\item
EURODEC - generator for high energy processes
\item
HERWIG - generator for hadron collisions
\item
ISAJET - the BNL generator for high energy proton-proton and
proton-antiproton collisions
\item
JETSET - Lund Monte Carlo for jet fragmentation and
e+e- anihilation
\item
PYTHIA - Lund Monte Carlo for high Pt hadron-hadron scattering
\item
LEPTO - Lund Monte Carlo for deep inelastic lepton-nucleon scattering
\item
PHOTOS - Monte Carlo for generation of radiative
corrections in decays
\item
PDFLIB - Parton Density Functions
\item
TWISTER - Monte Carlo for QCD high Pt scattering
\item
FRITIOF - hadron-hadron, hadron-nucleus, nucleus-nucleus interactions
\item
ARIADNE - Lund Monte Carlo for QCD cascades in the Colour Dipole approximation
\end{itemize}

\section{The CERNLIB modules}

The CERNLIB modules consist of complete standalone programs,
the most famous of which is surely \PAW{}. A brief
description of the various modules for a given system
is given in the system dependent installation section
of this manual.
