%.*****************************************************************>
%.* HPLOT  Reference Guide - Front Material                       *>
%.*  Last Mod.   11 May 1991  16.07  mg                           *>
%.*****************************************************************>
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    Tile page                                                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}
\large
\begin{center}CERN Program Library Long Writeup Y251\\[5cm]
{\Huge HPLOT}\\[1cm]
{\LARGE User's Guide}\\[2cm]
{\LARGE Version 5.40 (May 1991)}\\[5cm]
Ren\'e Brun\\[2mm]
Olivier Couet\\[2mm]
Nicole Cremel-Somon\\[1cm]
M. Goossens (Document Editor)\\[2cm]
CERN Geneva, Switzerland
\end{center}\end{titlepage}
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    Copyright  page                                               %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}
\framebox[\textwidth][t]{\hfill\begin{minipage}{0.92\textwidth}%
\vspace*{3mm}\begin{center}Copyright Notice\end{center}
\parskip\baselineskip
{\bf HPLOT -- User's Guide}
 
CERN Program Library entry {\bf Y251}
 
Copyright CERN, Geneva 1991
 
Copyright and any other appropriate legal protection of these
computer programs and associated documentation reserved in all
countries of the world.
 
These programs or documentation may not be reproduced by any
method without prior written consent of the Director-General
of CERN or his delegate.
 
Permission for the usage of any programs described herein is
granted apriori to those scientific institutes associated with
the CERN experimental program or with whom CERN has concluded
a scientific collaboration agreement.
 
Requests for information should be addressed to:
\begin{center}\begin{tabular}{l}
{\tt CERN Program Library Office}\\
{\tt CERN-CN Division            }\\
{\tt CH-1211 Geneva 23           }\\
{\tt Switzerland                 }\\
{\tt Tel.      +41 22 767 4951   }\\
{\tt Fax.      +41 22 767 7155   }\\
{\tt Bitnet:   CERNLIB@CERNVM    }\\
{\tt DECnet:   VXCERN::CERNLIB (node 22.190)  }\\
{\tt Internet: CERNLIB@CERNVM.CERN.CH}
\end{tabular}\end{center}\vspace*{2mm}
\end{minipage}\hfill}%end of minipage in framebox
\vspace{1cm}
\par {\bf Trademark notice: All trademarks appearing in this guide are acknowledged as such.}
\vfill
{\it Sixth  printing - May 1991}
\newpage
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    Introductory material                                         %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagenumbering{roman}
\setcounter{page}{1}
\section*{Preliminary remarks}
\par This {\bf user's guide} of the HPLOT package
consists of two parts:
\begin{OL}
\item An {\bf reference manual}, describing each available routine
in detail.
\item A set of {\bf complete example programs} with the output
they generate.
\end{OL}
\par The HPLOT program is implemented on various mainframes and personal
workstations. In particular versions exist for IBM~VM/CMS,
VAX/VMS and various Unix-like platforms, such as APOLLO, Cray,
Ultrix, IBM RS6000, Silicon Graphics and SUN.
\begin{center}
\fbox{\parbox{12cm}{Throughout this manual, commands to be {\bf entered}
are {\tt\underline{underlined}}}}
\end{center}
\index{underlining}
\index{user input}
\section*{Acknowledgements}
\par
The author is grateful to the many people who contributed to the HPLOT
project, through discussions or by providing code and assistance.
blabla blabla
\par This document has been produced using \LaTeX{} with
the {\tt CERNMAN} style option.
All pictures shown are produced with PAW and are included in
PostScript\cite{bib-POSTSC} format in the manual.
\section*{Related Documents}
\par This document can be complemented by the following documents:
\begin{UL}
\item PAW  - PAW blbla                       \cite{bib-PAW}
\end{UL}
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    Tables of contents ...                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents
\listoffigures
