%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                              %
%   PAW   - Reference Manual -- LaTeX Source                                   %
%                                                                              %
%   PAW system functions (used by kuipch2.tex)                                 %
%                                                                              %
%   Editor: Michel Goossens / CN-AS                                            %
%   Last Mod.: 6 October 1993 oc                                               %
%                                                                              %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Histograms inquiry functions}
\begin{UL}
\item\texttt{\$HEXIST(id)} returns 1 if histogram \texttt{id} exists or 0 otherwise
\item\texttt{\$HINFO(id,'ENTRIES')} returns the number of entries.
\item\texttt{\$HINFO(id,'MEAN')} returns the mean value.
\item\texttt{\$HINFO(id,'RMS')} returns the standard deviation.
\item\texttt{\$HINFO(id,'EVENTS')} returns the number of equivalent event.
\item\texttt{\$HINFO(id,'OVERFLOW')} returns the content of overflow channel.
\item\texttt{\$HINFO(id,'UNDERFLOW')} returns the content of underflow channel.
\item\texttt{\$HINFO(id,'MIN')} returns the minimum bin content.
\item\texttt{\$HINFO(id,'MAX')} returns the maximum bin content.
\item\texttt{\$HINFO(id,'SUM')} returns the total histogram content.
\item\texttt{\$HINFO(id,'XBINS')} returns the number of bins in X direction.
\item\texttt{\$HINFO(id,'XMIN')} returns the lower histogram limit in X direction.
\item\texttt{\$HINFO(id,'XMAX')} returns the upper histogram limit in X direction.
\item\texttt{\$HINFO(id,'YBINS')} returns the number of bins in Y direction.
\item\texttt{\$HINFO(id,'YMIN')} returns the lower histogram limit in Y direction.
\item\texttt{\$HINFO(id,'YMAX')} returns the upper histogram limit in Y direction.
\item\texttt{\$HTITLE(id)} returns the histogram title.
\end{UL}

\subsection{Graphics inquiry functions}
\begin{UL}
\item\texttt{\$GRAFINFO('XZONES')} returns the number of zones in X direction.
\item\texttt{\$GRAFINFO('YZONES')} returns the number of zones in Y direction.
\item\texttt{\$GRAFINFO('NT')} 
returns the current normalization transformation number.
\item\texttt{\$GRAFINFO('WNXMIN')} 
returns the lower X limit of window in current NT.
\item\texttt{\$GRAFINFO('WNXMAX')} 
returns the upper X limit of window in current NT.
\item\texttt{\$GRAFINFO('WNYMIN')} 
returns the lower Y limit of window in current NT.
\item\texttt{\$GRAFINFO('WNYMAX')} 
returns the upper Y limit of window in current NT.
\item\texttt{\$GRAFINFO('VPXMIN')} 
returns the lower X limit of viewport in current NT.
\item\texttt{\$GRAFINFO('VPXMAX')} 
returns the upper X limit of viewport in current NT.
\item\texttt{\$GRAFINFO('VPYMIN')} 
returns the lower Y limit of viewport in current NT.
\item\texttt{\$GRAFINFO('VPYMAX')} 
returns the upper Y limit of viewport in current NT.
\item\texttt{\$GRAFINFO('?attr')} returns the current value of the \HPLOT/\HIGZ{}
                              attribute \texttt{attr}. See the HELP of the command
                              \texttt{SET} to have the list of the valid values of
                              \texttt{attr}.
\end{UL}

\subsection{Cuts manipulations}
\begin{UL}
\item\texttt{\$CUT(n)} returns the cut expression \texttt{\$n}
\item\texttt{\$CUTEXPAND(string)} 
replace \texttt{\$n} in the (quoted) string by \texttt{\$CUT(n)}
\end{UL}
