\documentstyle[ifthen,calc,dingbat]{cernman}
\newcounter{Mycount}
\newcommand{\Denslist}
{\itemsep=0pt\parsep=0pt\partopsep=0pt\topsep=\baselineskip\parskip=0pt}
\newenvironment{EnumZW}{\renewcommand{\labelenumi}
                       {\setcounter{Mycount}{191+\value{enumi}}%
                       \raisebox{-2pt}[0cm][0cm]
                       {\Large\ding{\the\value{Mycount}}}}%
                       \enumerate\Denslist}%
                       {\endlist}
\newenvironment{EnumZB}{\renewcommand{\labelenumi}
                      {\setcounter{Mycount}{201+\value{enumi}}%
                   \raisebox{-2pt}[0cm][0cm]{\Large\ding{\the\value{Mycount}}}}%
                        \enumerate\Denslist}%
                       {\endlist}

%%%%%%%%% Specific commands for tagging PAW++ elements %%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\NbDW}[1]{\setcounter{Mycount}{191+#1}\ding{\the\value{Mycount}}}%
\newcommand{\NbDB}[1]{\setcounter{Mycount}{201+#1}\ding{\the\value{Mycount}}}%

\begin{document}

\NbDB{1} behind the application prompt. Via the
toggle buttons {H} \NbDB{4}

\begin{EnumZB}
\item In the } one can type, retrieve and edit command line with the help of
      a Korn shell emacs-style command line editing mode. See in appendix the 
      complete list of the editing keys.
\item The  \NbDB{2} shows the executed commands and command output.
      When in hold mode \NbDB{4} the transcript pad does not scroll to make 
      the new text
      visible. Mouse operations like ``Copy Paste'' are allowed in the 
      transcript pad. It is also possible to search a character string (see the
      menu bar description).
\item Every time the current directory is changed, the {\bf Current working 
      directory indicator} is updated. The current working directory can be 
\item Hold buttons.
\end{EnumZB}
\begin{EnumZW}
\item Allows manipulation of the 
\item Allows character string seach, copy/paste in the
\item Allows to invoke other panel. 
\item Some general settings are available in this menu. 
\item Online help.
\end{EnumZW}

\end{document}






